---
title: "Multidecadal Sea Ice Metrics Trend Analysis - Harmonized 12.5 km (OPTIMIZED)"
author: "Michael Wethington"
date: "Latest Version: 2025-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load_libraries, include=FALSE}
library(terra)
library(dplyr)
library(tidyr)
library(lubridate)
library(nlme)
library(ggplot2)
library(httr)
library(broom)
library(scales)
library(lmtest)
library(nortest)
library(changepoint)
library(RColorBrewer)
library(gridExtra)
library(readr)
```

# PART 1: DATA PREPARATION AND METRIC CALCULATION

## Calculate Regional Ice Extent Metrics

This section calculates daily sea ice concentration (SIC), sea ice extent, and SIC variability.

```{r calculate_regional_metrics}
calculate_stack_stats <- function(file_path) {
  raster_stack <- rast(file_path)
  layer_dates <- as.Date(as.numeric(time(raster_stack)), origin = "1970-01-01")
  
  # Apply threshold
  raster_stack <- app(raster_stack, fun = function(x) { x[x < 0.15] <- 0; return(x) })
  
  # Calculate metrics
  mean_sic <- global(raster_stack, fun = 'mean', na.rm = TRUE)[, 1]
  cell_area_sq_meters <- prod(res(raster_stack[[1]]))
  valid_ice_cells <- global(raster_stack >= 0.15, fun = 'sum', na.rm = TRUE)[, 1]
  total_ice_area_sq_km <- (valid_ice_cells * cell_area_sq_meters) / 1e6
  sic_sd <- global(raster_stack, fun = 'sd', na.rm = TRUE)[, 1]
  
  # Extract region name
  region_name <- tools::file_path_sans_ext(basename(file_path))
  region_name <- sub("NSIDC_25km_Harmonized_", "", region_name)
  
  return(data.frame(
    date = layer_dates, 
    mean_sic = mean_sic, 
    ice_extent_km = total_ice_area_sq_km, 
    sic_variability = sic_sd, 
    region = region_name
  ))
}

# Process all files
chunk_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/complete-harmonized-dataset/tif"
chunk_files <- list.files(chunk_dir, pattern = "\\.tif$", full.names = TRUE)

daily_df <- lapply(chunk_files, function(file) {
  cat("Processing file:", file, "\n")
  calculate_stack_stats(file)
}) %>% bind_rows()

# Clean and filter data
daily_df <- daily_df %>%
  mutate(
    region = ifelse(region == "nsidc_12_5km_harmonized_1979-2023", "Complete Study Area", region),
    Date = as.Date(date),
    Year = year(Date), 
    Month = month(Date), 
    Day = day(Date)
  ) %>%
  filter(Month %in% c(6, 7, 8, 9)) %>%
  rename(
    Region = region, 
    IceExtent_km = ice_extent_km, 
    MeanSIC = mean_sic, 
    SICVariability = sic_variability
  ) %>%
  select(-date)

# Calculate monthly variability
monthly_variability_df <- daily_df %>%
  group_by(Year, Month, Region) %>%
  summarize(
    MonthlySICVariability = sd(MeanSIC, na.rm = TRUE),
    MonthlyExtentVariability = sd(IceExtent_km, na.rm = TRUE),
    .groups = 'drop'
  )

# Save outputs
output_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/trend-analysis"
write.csv(daily_df, file.path(output_dir, "Harmonized_12km_Daily_Metrics.csv"), row.names = FALSE)
saveRDS(daily_df, file.path(output_dir, "Harmonized_12km_Daily_Metrics.rds"))
write.csv(monthly_variability_df, file.path(output_dir, "Harmonized_12km_Daily_Metrics_Monthly_Variability_Metrics.csv"), row.names = FALSE)
saveRDS(monthly_variability_df, file.path(output_dir, "Harmonized_12km_Daily_Metrics_Monthly_Variability_Metrics.rds"))
```


## Calculate Monthly Persistence and Duration Metrics

```{r calculate_persistence_duration}
# -------------------------------------------------------------------
# Updated sea-ice trend metrics to match Gentoo definitions
#   Duration  = mean length of all ice spells per pixel (SIC > threshold)
#   Persistence = mean open-water frequency per pixel (SIC < threshold)
# -------------------------------------------------------------------

library(terra)
library(dplyr)
library(lubridate)
library(purrr)

calculate_persistence_duration_stats <- function(file_path,
                                                 ice_threshold  = 0.15,
                                                 winter_months  = c(6, 7, 8, 9)) {
  # Read regional raster stack
  raster_stack <- rast(file_path)
  
  # Extract dates for each layer
  layer_dates <- as.Date(as.numeric(time(raster_stack)), origin = "1970-01-01")
  
  # Restrict to winter months
  winter_indices       <- which(month(layer_dates) %in% winter_months)
  winter_raster_stack  <- subset(raster_stack, winter_indices)
  winter_dates         <- layer_dates[winter_indices]
  
  # ------------------------------------------------------------------
  # Pixel-level helper functions (matching Gentoo model definitions)
  # ------------------------------------------------------------------
  
  # Mean duration of ice spells (SIC > threshold) for a single pixel
  calc_mean_duration <- function(x, threshold) {
    # x is the time series of SIC for a single pixel over the month
    if (all(is.na(x))) return(NA_real_)
    
    valid <- !is.na(x)
    if (!any(valid)) return(NA_real_)
    
    x_valid <- x[valid]
    rle_result <- rle(x_valid > threshold)
    
    # If no ice spells, duration = 0
    if (!any(rle_result$values)) return(0)
    
    mean(rle_result$lengths[rle_result$values])
  }
  
  # Open-water frequency (proportion of days with SIC < threshold) for a single pixel
  calc_open_water_freq <- function(x, threshold) {
    ow <- mean(x < threshold, na.rm = TRUE)
    if (is.nan(ow)) return(NA_real_)  # all NA
    ow
  }
  
  # ------------------------------------------------------------------
  # Aggregate by winter month
  # ------------------------------------------------------------------
  
  months <- unique(floor_date(winter_dates, "month"))
  
  results <- lapply(months, function(m) {
    month_indices       <- which(floor_date(winter_dates, "month") == m)
    month_raster_stack  <- subset(winter_raster_stack, month_indices)
    
    # Per-pixel mean ice-spell length over this month
    duration_rast <- app(
      month_raster_stack,
      fun = function(x) calc_mean_duration(x, ice_threshold)
    )
    
    # Per-pixel open-water frequency over this month
    open_water_rast <- app(
      month_raster_stack,
      fun = function(x) calc_open_water_freq(x, ice_threshold)
    )
    
    # Region-level summaries (mean and SD across pixels)
    mean_duration <- global(duration_rast,    fun = mean, na.rm = TRUE)[, 1]
    sd_duration   <- global(duration_rast,    fun = sd,   na.rm = TRUE)[, 1]
    mean_ow       <- global(open_water_rast,  fun = mean, na.rm = TRUE)[, 1]
    sd_ow         <- global(open_water_rast,  fun = sd,   na.rm = TRUE)[, 1]
    
    data.frame(
      Month                = m,
      # NOTE: Persistence here = open-water frequency (SIC < threshold)
      Persistence          = mean_ow,
      Persistence_SD       = sd_ow,
      # NOTE: Mean_Monthly_Duration here = mean length of all ice spells
      Mean_Monthly_Duration = mean_duration,
      Duration_SD          = sd_duration
    )
  }) %>% bind_rows()
  
  # Derive region name from file name
  region_name <- sub("NSIDC_25km_Harmonized_", "",
                     tools::file_path_sans_ext(basename(file_path)))
  results$Region <- region_name
  
  return(results)
}

# --------------------------------------------------------------------
# Process all regional files (chunks) using updated metric definitions
# --------------------------------------------------------------------

master_df <- lapply(chunk_files, function(file) {
  cat("Processing file:", file, "\n")
  calculate_persistence_duration_stats(file)
}) %>% bind_rows()

# --------------------------------------------------------------------
# Clean and export
# --------------------------------------------------------------------

cleaned_metrics <- master_df %>%
  rename(Date = Month) %>%
  mutate(
    Date   = as.Date(Date),
    Year   = year(Date),
    Month  = month(Date),
    Day    = day(Date),
    # Map full-region file to "Complete Study Area" label
    Region = ifelse(
      Region == "nsidc_12_5km_harmonized_1979-2023",
      "Complete Study Area",
      Region
    )
  ) %>%
  filter(Month %in% c(6, 7, 8, 9)) %>%  # winter months (redundant but explicit)
  distinct()

# Save outputs
write.csv(
  cleaned_metrics,
  file.path(output_dir, "Persistence_Duration_Metrics.csv"),
  row.names = FALSE
)

saveRDS(
  cleaned_metrics,
  file.path(output_dir, "Persistence_Duration_Metrics.rds")
)

```

# PART 2: CHANGEPOINT ANALYSIS

```{r changepoint_analysis}
# Load data
nsidc_metrics <- readRDS(file.path(output_dir, "Harmonized_12km_Daily_Metrics.rds"))

# Aggregate annual maximum extent
annual_max_extent <- nsidc_metrics %>%
  group_by(Year) %>%
  summarize(Max_Extent = min(IceExtent_km, na.rm = TRUE))

# Run changepoint analysis
cpt_annual_max <- cpt.mean(
  data = annual_max_extent$Max_Extent, 
  penalty = "BIC", 
  method = "SegNeigh", 
  Q = 15, 
  test.stat = "Normal", 
  minseglen = 1
)

# Plot results
plot(cpt_annual_max, main = "Changepoint Analysis of Maximum Annual Sea Ice Extent")
abline(v = annual_max_extent$Year[cpts(cpt_annual_max)], col = "red", lty = 2)

# Print changepoints
changepoints_years <- annual_max_extent$Year[cpts(cpt_annual_max)]
cat("Detected changepoints at years:", changepoints_years, "\n")
```

# PART 3: UNIFIED GLS ANALYSIS FUNCTION
GLS script:


This single function handles all metrics and time periods.

```{r unified_gls_function}
gls_analysis_function <- function(input_dataset, variable, output_directory, 
                                 start_year = NULL, end_year = NULL) {
  
  # Load dataset
  nsidc_metrics <- readRDS(input_dataset)
  
  # Ensure temporal columns exist
  if ("Date" %in% colnames(nsidc_metrics)) {
    nsidc_metrics <- nsidc_metrics %>%
      mutate(Year = year(Date), Month = month(Date), Day = day(Date))
  } else if ("Year" %in% colnames(nsidc_metrics) & "Month" %in% colnames(nsidc_metrics)) {
    nsidc_metrics <- nsidc_metrics %>% mutate(Day = 1)
  } else {
    stop("Dataset must contain either a 'Date' column or 'Year' and 'Month' columns.")
  }
  
  # Filter for winter months and year range
  winter_metrics <- nsidc_metrics %>%
    filter(Month %in% c(6, 7, 8, 9))
  
  if (!is.null(start_year)) winter_metrics <- winter_metrics %>% filter(Year >= start_year)
  if (!is.null(end_year)) winter_metrics <- winter_metrics %>% filter(Year <= end_year)
  
  # Calculate monthly averages
  monthly_averages <- winter_metrics %>%
    group_by(Year, Month, Region) %>%
    summarise(mean_var = mean(get(variable), na.rm = TRUE), .groups = 'drop') %>%
    ungroup() %>%
    mutate(
      Time = as.numeric(Year) + (Month - 1) / 12,
      month2 = month(ymd(paste(Year, Month, 1)), label = TRUE, abbr = FALSE)
    )
  
  # Fit GLS model function
  fit_gls_model <- function(region_data) {
    gls(mean_var ~ Time, correlation = corAR1(), data = region_data)
  }
  
  # Apply models for each region and month
  regions <- unique(monthly_averages$Region)
  months <- unique(monthly_averages$Month)
  gls_results <- list()
  
  for (region in regions) {
    for (month in months) {
      region_month_data <- monthly_averages %>% filter(Region == region, Month == month)
      if (nrow(region_month_data) > 1) {
        gls_results[[paste(region, month, sep = "_")]] <- fit_gls_model(region_month_data)
      }
    }
  }
  
  # Calculate annual metrics
  annual_metrics <- winter_metrics %>%
    group_by(Year, Region) %>%
    summarise(mean_var = mean(get(variable), na.rm = TRUE), .groups = 'drop') %>%
    ungroup() %>%
    mutate(Time = as.numeric(Year))
  
  for (region in regions) {
    region_annual_data <- annual_metrics %>% filter(Region == region)
    if (nrow(region_annual_data) > 1) {
      gls_results[[paste(region, "Annual", sep = "_")]] <- fit_gls_model(region_annual_data)
    }
  }
  
  # Extract and organize results
  num_tests <- length(gls_results)
  gls_summary_df <- lapply(names(gls_results), function(key) {
    model <- gls_results[[key]]
    summary_model <- summary(model)
    coefficients <- summary_model$tTable
    
    key_split <- unlist(strsplit(key, "_"))
    region <- key_split[1]
    month <- ifelse(length(key_split) > 1, key_split[2], "Annual")
    
    data.frame(
      Region = region,
      Metric = variable,
      Month = month,
      Intercept_Estimate = coefficients["(Intercept)", "Value"],
      Intercept_StdError = coefficients["(Intercept)", "Std.Error"],
      Intercept_tValue = coefficients["(Intercept)", "t-value"],
      Intercept_pValue = coefficients["(Intercept)", "p-value"],
      Time_Estimate = coefficients["Time", "Value"],
      Time_StdError = coefficients["Time", "Std.Error"],
      Time_tValue = coefficients["Time", "t-value"],
      Time_pValue = coefficients["Time", "p-value"],
      Adjusted_Time_pValue = p.adjust(coefficients["Time", "p-value"], method = "bonferroni", n = num_tests),
      AIC = AIC(model),
      BIC = BIC(model),
      LogLik = logLik(model),
      Residual_StdError = summary_model$sigma,
      DF_Residual = summary_model$dims$N - summary_model$dims$p,
      Start_Year = start_year,
      End_Year = end_year
    )
  }) %>% bind_rows()
  
  # Save results
  year_range <- paste(start_year, end_year, sep = "-")
  output_file <- file.path(output_directory, paste0("GLS_Summary_Results_with_Bonferroni_", variable, "_", year_range, ".csv"))
  write.csv(gls_summary_df, output_file, row.names = FALSE)
  
  return(gls_summary_df)
}
```

## Deploy GLS Analysis for All Metrics

```{r deploy_gls_analysis}
# Define analysis parameters
metrics_daily <- c("IceExtent_km", "MeanSIC", "SICVariability")
metrics_monthly_var <- c("MonthlySICVariability", "MonthlyExtentVariability")
metrics_persistence <- c("Persistence", "Persistence_SD", "Mean_Monthly_Duration", "Duration_SD")

year_ranges <- list(
  c(1979, 2024),
  c(2010, 2024)
)

input_files <- list(
  daily = file.path(output_dir, "Harmonized_12km_Daily_Metrics.rds"),
  monthly_var = file.path(output_dir, "Harmonized_12km_Daily_Metrics_Monthly_Variability_Metrics.rds"),
  persistence = file.path(output_dir, "Persistence_Duration_Metrics.rds")
)

trends_dir <- file.path(output_dir, "Trends")
dir.create(trends_dir, showWarnings = FALSE, recursive = TRUE)

# Run all analyses
all_analyses <- expand.grid(
  dataset = names(input_files),
  year_range_idx = seq_along(year_ranges),
  stringsAsFactors = FALSE
)

for (i in seq_len(nrow(all_analyses))) {
  dataset_type <- all_analyses$dataset[i]
  yr_idx <- all_analyses$year_range_idx[i]
  
  metrics <- switch(dataset_type,
    daily = metrics_daily,
    monthly_var = metrics_monthly_var,
    persistence = metrics_persistence
  )
  
  for (metric in metrics) {
    cat("Analyzing:", metric, "for years", year_ranges[[yr_idx]][1], "-", year_ranges[[yr_idx]][2], "\n")
    
    result <- gls_analysis_function(
      input_dataset = input_files[[dataset_type]],
      variable = metric,
      output_directory = trends_dir,
      start_year = year_ranges[[yr_idx]][1],
      end_year = year_ranges[[yr_idx]][2]
    )
  }
}
```

# PART 4: EXTRACT AND COMPILE SIGNIFICANT RESULTS

```{r extract_significant_results}
extract_and_compile_significant_results <- function(input_directory, output_file, significance_level = 0.05) {
  
  csv_files <- list.files(path = input_directory, pattern = "\\.csv$", full.names = TRUE)
  
  significant_results <- lapply(csv_files, function(file) {
    data <- read.csv(file)
    data %>%
      filter(Adjusted_Time_pValue < significance_level) %>%
      mutate(
        Intercept_CI_Lower = Intercept_Estimate - 1.96 * Intercept_StdError,
        Intercept_CI_Upper = Intercept_Estimate + 1.96 * Intercept_StdError,
        Time_CI_Lower = Time_Estimate - 1.96 * Time_StdError,
        Time_CI_Upper = Time_Estimate + 1.96 * Time_StdError
      )
  }) %>% bind_rows()
  
  write.csv(significant_results, output_file, row.names = FALSE)
  return(significant_results)
}

# Extract significant results for each time period
for (period in c("1979-2024", "2010-2024")) {
  period_dir <- file.path(trends_dir, period)
  dir.create(period_dir, showWarnings = FALSE)
  
  output_file <- file.path(period_dir, "significant_results_with_CI.csv")
  significant_results <- extract_and_compile_significant_results(period_dir, output_file)
  
  cat("Extracted", nrow(significant_results), "significant results for", period, "\n")
}
```

# PART 5: SOUTHERN OSCILLATION INDEX ANALYSIS

```{r soi_analysis}
# Load sea ice data
nsidc_metrics <- readRDS(file.path(output_dir, "Harmonized_12km_Daily_Metrics.rds"))

winter_metrics <- nsidc_metrics %>%
  mutate(Year = year(Date), Month = month(Date)) %>%
  filter(Month %in% c(6, 7, 8, 9))

monthly_averages <- winter_metrics %>%
  group_by(Year, Month, Region) %>%
  summarise(
    mean_extent = mean(IceExtent_km, na.rm = TRUE), 
    sd_extent = sd(IceExtent_km, na.rm = TRUE), 
    .groups = 'drop'
  ) %>%
  mutate(Time = as.numeric(Year) + (Month - 1) / 12)

# Load SOI data
url <- "https://www.cpc.ncep.noaa.gov/data/indices/soi"
soi_raw <- GET(url)
soi_text <- content(soi_raw, "text")
soi_lines <- strsplit(soi_text, "\n")[[1]]

standardized_soi_start <- grep("STANDARDIZED    DATA", soi_lines)
soi_standardized_lines <- soi_lines[(standardized_soi_start + 3):length(soi_lines)]
soi_standardized_lines <- soi_standardized_lines[!grepl("-999.9", soi_standardized_lines)]

soi_data <- read.table(text = soi_standardized_lines, fill = TRUE, stringsAsFactors = FALSE)
colnames(soi_data) <- c("Year", month.abb)

soi_data_long <- soi_data %>%
  mutate(Year = as.integer(Year)) %>%
  pivot_longer(-Year, names_to = "Month", values_to = "SOI") %>%
  mutate(
    Month = match(Month, month.abb), 
    Year = as.integer(Year),
    SOI = as.numeric(SOI)
  ) %>%
  filter(!is.na(SOI))

# Merge and calculate correlations
merged_data <- monthly_averages %>%
  inner_join(soi_data_long, by = c("Year", "Month")) %>%
  distinct()

correlation_results <- merged_data %>%
  group_by(Region, Month) %>%
  summarise(
    correlation = cor(mean_extent, SOI, use = "complete.obs"),
    cor_test = list(cor.test(mean_extent, SOI)),
    p.value = cor_test[[1]]$p.value,
    conf.low = cor_test[[1]]$conf.int[1],
    conf.high = cor_test[[1]]$conf.int[2],
    .groups = 'drop'
  )

# Visualize correlations
ggplot(correlation_results, aes(x = factor(Month, levels = 6:9), y = correlation, fill = Region)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, position = position_dodge(0.9)) +
  geom_text(aes(label = round(correlation, 2)), vjust = -1.5, position = position_dodge(0.9)) +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title = "Correlation between SOI and Sea Ice Extent by Region and Month",
    x = "Month", 
    y = "Correlation"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")

# Linear regression analysis
regression_results <- merged_data %>%
  group_by(Region, Month) %>%
  do(tidy(lm(mean_extent ~ SOI, data = .)))

soi_coefficients <- regression_results %>% filter(term == "SOI")

# Plot SOI effects
ggplot(soi_coefficients, aes(x = factor(Month), y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  facet_wrap(~ Region, scales = "free_y") +
  labs(
    title = "Effect of SOI on Sea Ice Extent Across Regions and Months",
    x = "Month",
    y = "SOI Coefficient Estimate"
  ) +
  theme_minimal()
```

# PART 6: TREND VISUALIZATION

```{r trend_heatmap}
# Load significant results
data <- read_csv(file.path(trends_dir, "2010-2024", "significant_results_with_CI.csv"))

# Clean and prepare data
data <- data %>%
  na.omit() %>%
  mutate(
    Metric = factor(Metric, levels = c(
      "Extent", "Extent (Var.)", "SIC", "SIC (Var.)", 
      "Duration", "Duration (Var.)", "Open Water Frequency", "Open Water Frequency (Var.)"
    )),
    Region = factor(Region, levels = c(
      "Complete Study Area", "Offshore", "Northern Shelf", "Middle Shelf", "Southern Shelf"
    )),
    Month = as.character(Month),
    Month = ifelse(Month == "Annual", "Winter", Month)
  )

# Create month labels
month_labels <- c("6" = "June", "7" = "July", "8" = "August", "9" = "September", "Winter" = "Winter")
data$Month <- factor(data$Month, levels = names(month_labels), labels = month_labels)

# Relative scaling function
relative_scale <- function(x) {
  if (min(x) == max(x)) return(rep(0, length(x)))
  rescale(abs(x), to = c(0, 1)) * sign(x)
}

data <- data %>%
  group_by(Metric) %>%
  mutate(Relative_Scaled_Slope = relative_scale(Time_Estimate)) %>%
  ungroup() %>%
  mutate(
    Significant = (Adjusted_Time_pValue <= 0.05),
    label = ifelse(Relative_Scaled_Slope > 0, "+", ifelse(Relative_Scaled_Slope < 0, "−", ""))
  ) %>%
  filter(Significant | abs(round(Relative_Scaled_Slope, 2)) > 0, Relative_Scaled_Slope != 0)

# Create heatmap
heatmap_plot <- ggplot(data, aes(x = Month, y = Region, fill = Relative_Scaled_Slope)) +
  geom_tile(color = "white") +
  geom_tile(data = filter(data, Significant), fill = NA, color = "black", size = 0.2) +
  geom_text(aes(label = label), size = 3, color = "black") +
  scale_fill_gradientn(
    colors = c("#E46726", "white", "#6D9EC1"), 
    name = "Trend Strength & Direction",
    limits = c(-1, 1),
    oob = squish
  ) +
  labs(title = "Seasonal Trend Directions and Strength by Region") +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.spacing = unit(2, "lines"),
    plot.margin = margin(10, 10, 10, 10)
  ) +
  facet_wrap(~ Metric, ncol = 2) +
  coord_fixed(ratio = 1, clip = "off")

print(heatmap_plot)

# Save plot
ggsave(
  filename = file.path(output_dir, "heatmap_plot.pdf"), 
  plot = heatmap_plot, 
  width = 10, 
  height = 8, 
  device = cairo_pdf
)
```

# PART 7: SPATIAL RASTER ANALYSIS

This function performs pixel-by-pixel trend analysis on raster stacks.

```{r spatial_analysis_function}
analyze_sea_ice_spatial <- function(start_year, end_year, use_mask = TRUE, 
                                   report_metrics = c("concentration", "duration", "persistence")) {
  
  # Define paths
  chunk_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/complete-harmonized-dataset/tif"
  chunk_files <- list.files(chunk_dir, pattern = "\\.tif$", full.names = TRUE)
  study_area_shapefile <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/gis-layers/study-area/shp/Frostbound_Study_Areas_EPSG_3976.shp"
  
  if (use_mask) {
    study_area <- vect(study_area_shapefile)
  }
  
  ice_threshold <- 0.15
  
  # Duration calculation function
  calculate_duration <- function(x, threshold) {
    if (all(is.na(x))) return(NA)
    rle_result <- rle(x > threshold)
    ifelse(any(rle_result$values), max(rle_result$lengths[rle_result$values]), 0)
  }
  
  # Process each chunk file
  process_chunk <- function(file_path) {
    message("Processing file: ", file_path)
    nsidc <- rast(file_path)
    
    annual_persistence <- list()
    annual_duration <- list()
    annual_concentration <- list()
    
    for (year in start_year:end_year) {
      message("Processing year: ", year)
      start_date <- as.Date(paste0(year, "-01-01"))
      end_date <- as.Date(paste0(year, "-12-31"))
      
      nsidc_year <- subset(nsidc, which(time(nsidc) >= start_date & time(nsidc) <= end_date))
      if (nlyr(nsidc_year) == 0) next
      
      if ("persistence" %in% report_metrics) {
        persistence <- app(nsidc_year, function(x) sum(ifelse(x < ice_threshold, 0, 1), na.rm = FALSE))
        names(persistence) <- paste0("Persistence_", year)
        annual_persistence[[as.character(year)]] <- persistence
      }
      
      if ("duration" %in% report_metrics) {
        duration <- app(nsidc_year, function(x) calculate_duration(x, ice_threshold))
        names(duration) <- paste0("Duration_", year)
        annual_duration[[as.character(year)]] <- duration
      }
      
      if ("concentration" %in% report_metrics) {
        concentration <- app(nsidc_year, function(x) mean(x, na.rm = TRUE))
        names(concentration) <- paste0("Concentration_", year)
        annual_concentration[[as.character(year)]] <- concentration
      }
    }
    
    return(list(persistence = annual_persistence, duration = annual_duration, concentration = annual_concentration))
  }
  
  # Process all files
  results <- lapply(chunk_files, process_chunk)
  
  # Combine rasters
  combine_rasters <- function(raster_list) {
    raster_stack <- rast()
    for (year in names(raster_list)) {
      if (!is.null(raster_list[[year]])) {
        raster_stack <- c(raster_stack, raster_list[[year]])
      }
    }
    return(raster_stack)
  }
  
  persistence_stack <- if ("persistence" %in% report_metrics) {
    stack <- combine_rasters(do.call(c, lapply(results, `[[`, "persistence")))
    if (use_mask) mask(stack, study_area) else stack
  } else NULL
  
  duration_stack <- if ("duration" %in% report_metrics) {
    stack <- combine_rasters(do.call(c, lapply(results, `[[`, "duration")))
    if (use_mask) mask(stack, study_area) else stack
  } else NULL
  
  concentration_stack <- if ("concentration" %in% report_metrics) {
    stack <- combine_rasters(do.call(c, lapply(results, `[[`, "concentration")))
    if (use_mask) mask(stack, study_area) else stack
  } else NULL
  
  # Apply linear model to each pixel
  apply_lm <- function(y) {
    if (all(is.na(y))) return(c(NA, NA))
    years <- start_year:end_year
    fit <- tryCatch(lm(y ~ years), error = function(e) return(NULL))
    if (is.null(fit)) return(c(NA, NA))
    c(coef(fit)[2], summary(fit)$coefficients[2, 4])
  }
  
  # Calculate trends
  persistence_trend <- if (!is.null(persistence_stack)) app(persistence_stack, apply_lm) else NULL
  duration_trend <- if (!is.null(duration_stack)) app(duration_stack, apply_lm) else NULL
  concentration_trend <- if (!is.null(concentration_stack)) app(concentration_stack, apply_lm) else NULL
  
  # Extract slopes and p-values
  extract_layers <- function(trend) {
    if (is.null(trend)) return(list(slope = NULL, pvalue = NULL))
    list(slope = trend[[1]], pvalue = trend[[2]])
  }
  
  persistence_layers <- extract_layers(persistence_trend)
  duration_layers <- extract_layers(duration_trend)
  concentration_layers <- extract_layers(concentration_trend)
  
  # Calculate and print statistics
  calculate_stats <- function(slope, pvalue, metric_name) {
    if (is.null(slope)) return(NULL)
    
    values_slope <- values(slope)
    values_pvalue <- values(pvalue)
    valid_data <- cbind(values_slope, values_pvalue)
    valid_data <- valid_data[complete.cases(valid_data), ]
    
    alpha <- 0.05
    stats <- data.frame(
      Metric = c("Mean slope", "Median slope", "Min slope", "Max slope", 
                 "% Significant", "% Positive", "% Negative", "% Non-significant"),
      Value = c(
        mean(valid_data[, 1]),
        median(valid_data[, 1]),
        min(valid_data[, 1]),
        max(valid_data[, 1]),
        sum(valid_data[, 2] <= alpha) / nrow(valid_data) * 100,
        sum(valid_data[, 2] <= alpha & valid_data[, 1] > 0) / nrow(valid_data) * 100,
        sum(valid_data[, 2] <= alpha & valid_data[, 1] < 0) / nrow(valid_data) * 100,
        100 - sum(valid_data[, 2] <= alpha) / nrow(valid_data) * 100
      )
    )
    names(stats)[2] <- metric_name
    print(stats)
    return(stats)
  }
  
  calculate_stats(persistence_layers$slope, persistence_layers$pvalue, "Persistence")
  calculate_stats(duration_layers$slope, duration_layers$pvalue, "Duration")
  calculate_stats(concentration_layers$slope, concentration_layers$pvalue, "Concentration")
  
  # Plot results
  trend_colors <- colorRampPalette(c("red", "white", "blue"))
  alpha <- 0.05
  
  par(mfrow = c(2, 3), mar = c(4, 4, 2, 1))
  
  plot_trend <- function(slope, pvalue, title) {
    if (is.null(slope)) return()
    plot(slope, col = trend_colors(100), main = title, legend = TRUE, colNA = "black")
  }
  
  plot_significance <- function(slope, pvalue, title) {
    if (is.null(pvalue)) return()
    significant_cells <- pvalue <= alpha
    significant_mask <- ifel(significant_cells, 1, NA)
    positive_trend <- ifel(significant_mask == 1 & slope > 0, 1, NA)
    negative_trend <- ifel(significant_mask == 1 & slope < 0, 2, NA)
    na_mask <- ifel(is.na(slope), 3, NA)
    
    combined_trend <- positive_trend
    combined_trend[!is.na(negative_trend)] <- 2
    combined_trend[!is.na(na_mask)] <- 3
    
    color_table <- data.frame(value = c(1, 2, 3), col = c("blue", "red", "black"))
    coltab(combined_trend) <- color_table
    plot(combined_trend, main = title)
    legend("topright", legend = c("Positive", "Negative", "NA"), fill = color_table$col)
  }
  
  plot_trend(persistence_layers$slope, persistence_layers$pvalue, "Persistence Trend")
  plot_trend(duration_layers$slope, duration_layers$pvalue, "Duration Trend")
  plot_trend(concentration_layers$slope, concentration_layers$pvalue, "Concentration Trend")
  
  plot_significance(persistence_layers$slope, persistence_layers$pvalue, "Persistence Significance")
  plot_significance(duration_layers$slope, duration_layers$pvalue, "Duration Significance")
  plot_significance(concentration_layers$slope, concentration_layers$pvalue, "Concentration Significance")
}

# Example usage (uncomment to run)
# analyze_sea_ice_spatial(start_year = 1981, end_year = 2023, use_mask = FALSE, 
#                        report_metrics = c("concentration", "duration", "persistence"))
```


