---
title: "NSIDC_Sea-Ice-Index_Processing Pipeline"
author: "Michael Wethington"
date: "2024-04-21"
output: html_document
---

Source Data Concentration Surface Classification Values:

# SIC: 0-1000
# Pole hole: 2510
# Coast line: 2530
# Land: 2540
# Missing 2550


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Reproject the downloaded tifs to EPSG:3976**

    1. Generate a list of all the downloaded NSDIC 25.5 km tifs
    2. Loop through list of raster and ensure that the projection is EPSG: 3976
    3. Export each raster 

```{r include=FALSE}
# Load required libraries
library(terra)   # For spatial data analysis and raster manipulation

# Define directories
input_raster_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/staged"
output_raster_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/processed"

# Ensure output directory exists
if (!dir.exists(output_raster_dir)) {
  dir.create(output_raster_dir, recursive = TRUE)
}

# List all raster files in the input directory
raster_files <- list.files(path = input_raster_dir, pattern = "\\.tif$", full.names = TRUE)


# Process each raster file
for (file in raster_files) {
  r <- rast(file)  # Load the raster file
  

  # Reproject the raster to EPSG 3976
  r_reprojected <- project(r, crs("EPSG:3976"))

  # Define the output filename using the original file's name
  output_filename <- file.path(output_raster_dir, basename(file))
  
  # Save the reprojected raster to the output directory
  writeRaster(r_reprojected, filename=output_filename, overwrite=TRUE)
  
  cat("Reprojected to EPSG:3976 and saved:", output_filename, "\n")
}
```


**Pull (Clip) raster masks of the study area**


```{r}
# Load required libraries
library(terra)   # For spatial data analysis and raster manipulation

# Define directories
input_raster_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/processed"

output_raster_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/processed_clipped"

# Ensure output directory exists
if (!dir.exists(output_raster_dir)) {
  dir.create(output_raster_dir, recursive = TRUE)
}

# Load the study area shapefile
study_area_shp <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/gis-layers/study-area/Frostbound_AQ_Subregions_EPSG_3976.shp"
study_area <- vect(study_area_shp)

# List all raster files in the input directory and sort them by date
raster_files <- list.files(path = input_raster_dir, pattern = "\\.tif$", full.names = TRUE)

# Function to extract date from filename for sorting
extractDate <- function(filename) {
  # Regex to extract date assuming format YYYYMMDD in filename
  dates <- as.Date(sub(".*(\\d{4})(\\d{2})(\\d{2}).*", "\\1-\\2-\\3", basename(filename)), format="%Y-%m-%d")
  return(dates)
}

# Sort files by date
raster_files <- raster_files[order(sapply(raster_files, extractDate))]

# Process each raster file
for (file in raster_files) {
  r <- rast(file)
  # Crop raster to the extent of the study area
  r_cropped <- crop(r, study_area)
  
  # Define the output filename using the original file's name
  output_filename <- file.path(output_raster_dir, basename(file))
  
  # Save the cropped raster to the output directory
  writeRaster(r_cropped, filename=output_filename, overwrite=TRUE)
  
  cat("Processed and saved:", output_filename, "\n")
}

```


**Generate a netcdf Raster Stack for the dataset**


```{r} 

library(terra)

# Function to extract the date from a filename
extractDate <- function(filename) {
  base_name <- basename(filename)
  date_string <- sub("S_(\\d{8})_.*", "\\1", base_name)
  date_converted <- as.Date(date_string, format="%Y%m%d")
  if (is.na(date_converted)) {
    warning("Date conversion failed for: ", base_name)
  }
  return(date_converted)
}

# Define the directory with raster files
raster_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/processed_clipped"

raster_files <- list.files(path = raster_dir, pattern = "\\.tif$", full.names = TRUE)
raster_files <- raster_files[order(sapply(raster_files, extractDate))]

# Create a raster stack
raster_stack <- rast(raster_files)

# Naming layers by date extracted from filenames
names(raster_stack) <- sapply(raster_files, function(x) format(extractDate(x), "%Y-%m-%d"))

# Assigning time values
dates <- as.Date(sapply(raster_files, extractDate))
time(raster_stack) <- dates

# Set values outside 0-1000 and specific feature values to NA
# Pole hole: 2510
# Coast line: 2530
# Land: 2540
# Missing 2550

values_to_exclude <- c(2510, 2530, 2540, 2550)  # Other non-ice values

raster_stack[raster_stack %in% values_to_exclude] <- NA


# Scale the values from 0-1000 to 0-100% for percentage representation
raster_stack <- raster_stack / 10


# Check and print time data for verification
if (length(time(raster_stack)) > 0 && !all(is.na(time(raster_stack)))) {
  print("Time data are present and appear correctly set.")
  print(head(time(raster_stack), 20))  # Print the first 20 time entries to verify
} else {
  print("No valid time data found or all entries are NA.")
}

# Output file path for NetCDF
nc_output_path <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/NSIDC_25km_Annual_Study-Area.nc"

# Write the processed raster stack to a NetCDF file
# writeCDF(raster_stack, filename = nc_output_path, overwrite = TRUE, varname="sea_ice_concentration", 
#          longname="Sea Ice Concentration")

writeCDF(raster_stack, filename = nc_output_path, overwrite = TRUE)

# # Load and print the saved raster stack to verify
loaded_stack <- rast(nc_output_path)
print(loaded_stack)
head(time(loaded_stack))
# 
# # Check and print time data for verification
# if (length(time(loaded_stack)) > 0 && !all(is.na(time(loaded_stack)))) {
#   print("Time data are present and appear correctly set.")
#   print(head(time(loaded_stack), 20))  # Print the first 20 time entries to verify
# } else {
#   print("No valid time data found or all entries are NA.")
# }

```


**Create a Substack for Antarctic Winter Months**

    1. Subset the data for June-September
    2. Generate and export a new winter raster stack 
    
    
```{r Generate Winter Substack}
library(terra)    # For raster data manipulation
library(lubridate)  # For date-time operations

# Load the complete raster stack
nsidc_full <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/NSIDC_25km_Annual_Study-Area.nc")

# # Extract dates from raster layer names or metadata
dates <- time(nsidc_full)
# 
# # Identify layers corresponding to the Antarctic winter months (June to September)
winter_indices <- which(month(dates) %in% 6:9)
# 
# # Create a substack with only the winter month layers
nsidc_winter <- subset(nsidc_full, winter_indices)

# Extract dates from raster layer names or metadata
winter_dates <- time(nsidc_winter)

# Extract unique months from the dates
unique_months <- unique(month(winter_dates))

# Print the unique months to confirm they are June, July, August, and September
print(unique_months)
# 
# # save this substack 
writeCDF(nsidc_winter, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/NSIDC_25km_Winter_Months.nc", overwrite=TRUE)

#Check to make sure it loads
nsidc_winter <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/NSIDC_25km_Winter_Months.nc")

```

**Create Subregion Stacks**

    1. Loop through the individual study area subregions (shapefile)
    2. Generate and export new raster stacks for each respective subregion


```{r}
library(terra)

# Load the raster stack
nsidc <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/NSIDC_25km_Winter_Months.nc")
# Load the study area shapefile
study_areas <- vect("D:/Manuscripts_localData/FrostBound_AQ/Datasets/gis-layers/study-area/Frostbound_AQ_Subregions_EPSG_3976.shp")

# Define the output directory
output_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/substack/"

# Function to create subsets and write to .nc files
create_subset_and_export <- function(region_name) {
  # Subset the shapefile for the given region
  region_shape <- study_areas[study_areas$Region == region_name, ]

  # Crop and mask the raster stack by the region shape
  region_raster <- crop(nsidc, region_shape)
  region_raster <- mask(region_raster, region_shape)

  # Define the output file name
  output_file <- paste0(output_dir, "NSIDC_25km_", gsub(" ", "_", region_name), ".nc")

  # Write the subset raster stack to a .nc file
  writeCDF(region_raster, output_file, overwrite = TRUE)
}

# Extract unique region names
regions <- unique(study_areas$Region)

# Create subsets and export for each region
for (region in regions) {
  create_subset_and_export(region)
}


# # Write the subset raster stack to a .nc file
writeCDF(nsidc, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/substack/NSIDC_25km_Full_Study_Area.nc", overwrite = TRUE)




```



