---
title: "Sea Ice Trend Analysis - NSIDC"
author: "Michael Wethington"
date: "2024-05-14"
output: html_document
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(terra)

```

## **Dataset Overview**

```{r Data Overview}

nsidc_winter <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/25km_Sea-Ice-Index/stack/substack/NSIDC_25km_Full_Study_Area.nc"

nsidc_winter <- rast(nsidc_winter)

# Print the object to see its structure
print(nsidc_winter)

# Summary of the object
summary(nsidc_winter)

# If it's a SpatRaster, use terra functions to explore further
if (inherits(nsidc_winter, "SpatRaster")) {
  # Check the number of layers
  nlyr(nsidc_winter)
  
  # Extract dates from raster layer names or metadata
  winter_dates <- time(nsidc_winter)

# Extract unique months from the dates
  unique_months <- unique(month(winter_dates))
  
  print(unique_months)
  
  # Get the names of the layers
  names(nsidc_winter)
  
  # Plot a single layer
  plot(nsidc_winter[[1]])
  
  # Get the extent of the raster
  ext(nsidc_winter)
  
  # Get the resolution of the raster
  res(nsidc_winter)
  
  # Check the CRS (Coordinate Reference System)
  crs(nsidc_winter)
} else {
  print("The object is not a SpatRaster. Further exploration is needed.")
}


summary(nsidc_winter[[1]])

plot(nsidc_winter[[1]])

rm(nsidc_winter)
```




**Calculate Regional Ice Extent Metrics (Parallel)**

```{r Calculate Regional Ice Extent Metrics}


library(terra)        # For raster data manipulation
library(doParallel)   # For parallel processing
library(foreach)      # For looping with parallel support
library(dplyr)
library(lubridate)

# Function to calculate mean SIC and sea ice extent for a given raster layer within a file
calculate_layer_stats <- function(layer_index, file_path) {
  raster_stack <- rast(file_path)                  # Load the raster stack from file
  layer_data <- raster_stack[[layer_index]]        # Extract the specific layer
  
  layer_date <- time(layer_data)                   # Extract date from layer metadata

  # Handle sea ice concentration: treat cells < 15 as 0
  layer_data[layer_data < 15] <- 0
  
  # Calculate mean SIC excluding NAs
  mean_sic <- as.numeric(terra::global(layer_data, fun = 'mean', na.rm = TRUE))  
  
  # Handle sea ice extent: count cells >= 15
  valid_ice_cells <- sum(values(layer_data) >= 15, na.rm = TRUE)
  cell_area_sq_meters <- prod(res(layer_data))    # Calculate the area of one cell in square meters
  total_ice_area_sq_km <- (valid_ice_cells * cell_area_sq_meters) / 1e6  # Convert total ice area to square kilometers
  
  rm(raster_stack, layer_data)   # Clear memory
  gc()                           # Run garbage collection
  
  # Extract region name from the file path
  region_name <- tools::file_path_sans_ext(basename(file_path))
  region_name <- sub("NSIDC_25km_", "", region_name)  # Adjust based on actual pattern in filenames
  
  return(list(mean_sic = mean_sic, ice_extent_km = total_ice_area_sq_km, date = layer_date, region = region_name))
}

# Directory with chunk files for each region
chunk_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/25km_Sea-Ice-Index/stack/substack"

chunk_files <- list.files(chunk_dir, pattern = "\\.nc$", full.names = TRUE)  # List all .nc files for each region


all_results <- list()  # Initialize list to store results from all chunks

for (r_fp in chunk_files) {
  numCores <- detectCores() - 1
  cl <- makeCluster(numCores)
  registerDoParallel(cl)
  
  clusterExport(cl, varlist = c("calculate_layer_stats", "r_fp"))
  
  results <- foreach(i = 1:nlyr(rast(r_fp)), .packages = 'terra', .errorhandling = "pass") %dopar% {
    tryCatch({
      calculate_layer_stats(i, r_fp)
    }, error = function(e) {
      return(list(error = TRUE, message = e$message))
    })
  }
  
  stopCluster(cl)
  
  # Append chunk results to overall results, add file path to results for region tracking
  all_results <- c(all_results, lapply(results, function(x) c(x, file = r_fp)))
}

# Convert list of all results into a dataframe, handling possible errors
final_result_df <- do.call(rbind, lapply(all_results, function(x) {
  if (!is.null(x$error)) {
    return(data.frame(Date = NA, MeanSIC = NA, IceExtent_km = NA, Region = NA, error = x$message))
  } else {
    return(data.frame(Date = x$date, MeanSIC = x$mean_sic, IceExtent_km = x$ice_extent_km, Region = x$region, error = NA))
  }
}))

# Add year, month, and day columns
final_result_df <- final_result_df %>%
  dplyr::mutate(Year = year(Date), Month = month(Date), Day = day(Date))

# Filter for Full_Study_Area and June, July, August, and September
final_result_df <- final_result_df %>%
  filter(Month %in% c(6, 7, 8, 9))

# Save combined results to CSV and RDS for further use
write.csv(final_result_df, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/25km_Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.csv", row.names = FALSE)
saveRDS(final_result_df, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/25km_Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.rds")



```


## **Calculate Sea Ice Persistence**


```{r Calculate Sea Ice Persistence}
library(terra)
library(lubridate)

# Load the NSIDC sea ice concentration data
nsidc <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/NSIDC_25km_Study-Area.nc")

# Define a function to count days with ice cover > 15% for each year
count_ice_days <- function(x) {
  ice_days <- sum(x > 15, na.rm = TRUE)
  return(ice_days)
}

# Extract the years from the time attribute using lubridate
years <- unique(year(time(nsidc)))

# Initialize an empty list to store the annual ice day count rasters
annual_ice_days <- list()

# Loop over each year, subset the raster stack, and apply the counting function
for (year in years) {
  year_rasters <- subset(nsidc, year(time(nsidc)) == year)
  # Apply count_ice_days function across layers (days) at each pixel
  ice_days_raster <- app(year_rasters, count_ice_days)  # No need for 'along=3', app inherently operates cell-wise
  names(ice_days_raster) <- paste("Ice_Days", year, sep="_")
  annual_ice_days[[year]] <- ice_days_raster
  print(paste("Processed year:", year))
}

# Combine the annual rasters into a single stack
annual_ice_days_stack <- rast(annual_ice_days)

# Save the resulting stack to disk
writeRaster(annual_ice_days_stack, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR-Unified_12km/stack/substack/Annual_Ice_Days_Stack.tif", overwrite=TRUE)


```


**Plot Sea Ice Persistence Trends**

```{r Plot Sea Ice Persistence Trends}

library(terra)
library(RColorBrewer)

# Assuming 'annual_ice_days_stack' is already loaded and contains the necessary layers

# Define the color palette
color_palette <- rev(brewer.pal(9, "Blues"))

# Set up the multi-panel plot layout
par(mfrow = c(2, 2))

# Plot each year's data using a consistent color scale from 0 to 365 days
plot(annual_ice_days_stack$Ice_Days_1988, main = "Ice Days 1988", col=color_palette, breaks=seq(0, 365, length.out = length(color_palette) + 1))
plot(annual_ice_days_stack$Ice_Days_1990, main = "Ice Days 1995", col=color_palette, breaks=seq(0, 365, length.out = length(color_palette) + 1))
plot(annual_ice_days_stack$Ice_Days_2020, main = "Ice Days 2020", col=color_palette, breaks=seq(0, 365, length.out = length(color_palette) + 1))
plot(annual_ice_days_stack$Ice_Days_2023, main = "Ice Days 2023", col=color_palette, breaks=seq(0, 365, length.out = length(color_palette) + 1))

# Reset the plot layout
par(mfrow = c(1, 1))

```


Full Region

```{r Full Region Annual Trend Visualization}
library(dplyr)
library(ggplot2)
library(lubridate)
library(wesanderson)
library(scales)

# Load your dataset
nsidc_metrics <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.rds")

# Add year, month, and day columns
nsidc_metrics <- nsidc_metrics %>%
  mutate(Year = year(Date), Month = month(Date), Day = day(Date))

# Filter for Full_Study_Area and June, July, August, and September
nsidc_metrics_filtered <- nsidc_metrics %>%
  filter(Region == "Full_Study_Area" & Month %in% c(6, 7, 8, 9))

# Add a month2 column with month names
nsidc_metrics_filtered <- nsidc_metrics_filtered %>%
  mutate(month2 = month(Date, label = TRUE, abbr = FALSE))

# Choose a light red color for the SD bars
sd_color <- "#7AB8BF"  # Light red color (you can adjust to your preferred shade)
axis_color <- "gray70"  # Light gray color for the axis lines

# Create the plot
p <- nsidc_metrics_filtered %>%
  group_by(month2, Year) %>%
  summarise(mean_extent = mean(IceExtent_km, na.rm = TRUE), 
            sd_extent = sd(IceExtent_km, na.rm = TRUE), .groups = 'drop') %>%
  ggplot(aes(x = Year, y = mean_extent)) +
  geom_errorbar(aes(ymin = mean_extent - sd_extent, ymax = mean_extent + sd_extent), 
                width = 0.2, color = sd_color) +
  geom_point(color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  facet_wrap(~month2, scales = "fixed", ncol = 3) +
  xlab(NULL) +
  ylab(bquote("Ice extent" ~ (km^2 %*% 10^6))) +
  scale_y_continuous(labels = scales::scientific) +
  labs(
    title = sprintf(
      "Study Area Wide: Sea ice extent between %d and %d",
      min(nsidc_metrics_filtered$Year),
      max(nsidc_metrics_filtered$Year)
    ),
    subtitle = "The vertical bar at each point shows the standard deviation around the mean. The blue line represents the linear trend.",
    caption = "Source: NSIDC 25km Sea Ice Index dataset"
  ) +
  theme(
    plot.caption = element_text(size = 8, color = "black"),
    plot.margin = unit(c(5.5, 10, 5.5, 5.5), "points"),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black", size = 12, face = "bold"),
    panel.grid = element_blank(),
    axis.line = element_line(color = axis_color),
    legend.position = "none"
  )

print(p)
# 
# # Save the filtered data
# write.csv(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.csv", row.names = FALSE)
# saveRDS(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.rds")



```


Northern Shelf

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(wesanderson)
library(scales)

# Load your dataset
nsidc_metrics <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.rds")

# Add year, month, and day columns
nsidc_metrics <- nsidc_metrics %>%
  mutate(Year = year(Date), Month = month(Date), Day = day(Date))

# Filter for Full_Study_Area and June, July, August, and September
nsidc_metrics_filtered <- nsidc_metrics %>%
  filter(Region == "Northern_Shelf_Region" & Month %in% c(6, 7, 8, 9))

# Add a month2 column with month names
nsidc_metrics_filtered <- nsidc_metrics_filtered %>%
  mutate(month2 = month(Date, label = TRUE, abbr = FALSE))

# Choose a light red color for the SD bars
sd_color <- "#7AB8BF"  # Light red color (you can adjust to your preferred shade)
axis_color <- "gray70"  # Light gray color for the axis lines

# Create the plot
p <- nsidc_metrics_filtered %>%
  group_by(month2, Year) %>%
  summarise(mean_extent = mean(IceExtent_km, na.rm = TRUE), 
            sd_extent = sd(IceExtent_km, na.rm = TRUE), .groups = 'drop') %>%
  ggplot(aes(x = Year, y = mean_extent)) +
  geom_errorbar(aes(ymin = mean_extent - sd_extent, ymax = mean_extent + sd_extent), 
                width = 0.2, color = sd_color) +
  geom_point(color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  facet_wrap(~month2, scales = "fixed", ncol = 3) +
  xlab(NULL) +
  ylab(bquote("Ice extent" ~ (km^2 %*% 10^6))) +
  scale_y_continuous(labels = scales::scientific) +
  labs(
    title = sprintf(
      "Northern Shelf Region: Sea ice extent between %d and %d",
      min(nsidc_metrics_filtered$Year),
      max(nsidc_metrics_filtered$Year)
    ),
    subtitle = "The vertical bar at each point shows the standard deviation around the mean. The blue line represents the linear trend.",
    caption = "Source: NSIDC 25km Sea Ice Index dataset"
  ) +
  theme(
    plot.caption = element_text(size = 8, color = "black"),
    plot.margin = unit(c(5.5, 10, 5.5, 5.5), "points"),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black", size = 12, face = "bold"),
    panel.grid = element_blank(),
    axis.line = element_line(color = axis_color),
    legend.position = "none"
  )

print(p)
# 
# # Save the filtered data
# write.csv(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.csv", row.names = FALSE)
# saveRDS(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.rds")

```


```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(wesanderson)
library(scales)

# Load your dataset
nsidc_metrics <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.rds")

# Add year, month, and day columns
nsidc_metrics <- nsidc_metrics %>%
  mutate(Year = year(Date), Month = month(Date), Day = day(Date))

# Filter for Full_Study_Area and June, July, August, and September
nsidc_metrics_filtered <- nsidc_metrics %>%
  filter(Region == "Mid-Shelf_Region" & Month %in% c(6, 7, 8, 9))

# Add a month2 column with month names
nsidc_metrics_filtered <- nsidc_metrics_filtered %>%
  mutate(month2 = month(Date, label = TRUE, abbr = FALSE))

# Choose a light red color for the SD bars
sd_color <- "#7AB8BF"  # Light red color (you can adjust to your preferred shade)
axis_color <- "gray70"  # Light gray color for the axis lines

# Create the plot
p <- nsidc_metrics_filtered %>%
  group_by(month2, Year) %>%
  summarise(mean_extent = mean(IceExtent_km, na.rm = TRUE), 
            sd_extent = sd(IceExtent_km, na.rm = TRUE), .groups = 'drop') %>%
  ggplot(aes(x = Year, y = mean_extent)) +
  geom_errorbar(aes(ymin = mean_extent - sd_extent, ymax = mean_extent + sd_extent), 
                width = 0.2, color = sd_color) +
  geom_point(color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  facet_wrap(~month2, scales = "fixed", ncol = 3) +
  xlab(NULL) +
  ylab(bquote("Ice extent" ~ (km^2 %*% 10^6))) +
  scale_y_continuous(labels = scales::scientific) +
  labs(
    title = sprintf(
      "Middle Shelf Region: Sea ice extent between %d and %d",
      min(nsidc_metrics_filtered$Year),
      max(nsidc_metrics_filtered$Year)
    ),
    subtitle = "The vertical bar at each point shows the standard deviation around the mean. The blue line represents the linear trend.",
    caption = "Source: NSIDC 25km Sea Ice Index dataset"
  ) +
  theme(
    plot.caption = element_text(size = 8, color = "black"),
    plot.margin = unit(c(5.5, 10, 5.5, 5.5), "points"),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black", size = 12, face = "bold"),
    panel.grid = element_blank(),
    axis.line = element_line(color = axis_color),
    legend.position = "none"
  )

print(p)
# 
# # Save the filtered data
# write.csv(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.csv", row.names = FALSE)
# saveRDS(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.rds")

```
Southern Shelf Region Trends
```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(wesanderson)
library(scales)

# Load your dataset
nsidc_metrics <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.rds")

# Add year, month, and day columns
nsidc_metrics <- nsidc_metrics %>%
  mutate(Year = year(Date), Month = month(Date), Day = day(Date))

# Filter for Full_Study_Area and June, July, August, and September
nsidc_metrics_filtered <- nsidc_metrics %>%
  filter(Region == "Southern_Shelf_Region" & Month %in% c(6, 7, 8, 9))

# Add a month2 column with month names
nsidc_metrics_filtered <- nsidc_metrics_filtered %>%
  mutate(month2 = month(Date, label = TRUE, abbr = FALSE))

# Choose a light red color for the SD bars
sd_color <- "#7AB8BF"  # Light red color (you can adjust to your preferred shade)
axis_color <- "gray70"  # Light gray color for the axis lines

# Create the plot
p <- nsidc_metrics_filtered %>%
  group_by(month2, Year) %>%
  summarise(mean_extent = mean(IceExtent_km, na.rm = TRUE), 
            sd_extent = sd(IceExtent_km, na.rm = TRUE), .groups = 'drop') %>%
  ggplot(aes(x = Year, y = mean_extent)) +
  geom_errorbar(aes(ymin = mean_extent - sd_extent, ymax = mean_extent + sd_extent), 
                width = 0.2, color = sd_color) +
  geom_point(color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  facet_wrap(~month2, scales = "fixed", ncol = 3) +
  xlab(NULL) +
  ylab(bquote("Ice extent" ~ (km^2 %*% 10^6))) +
  scale_y_continuous(labels = scales::scientific) +
  labs(
    title = sprintf(
      "Southern Shelf Region: Sea ice extent between %d and %d",
      min(nsidc_metrics_filtered$Year),
      max(nsidc_metrics_filtered$Year)
    ),
    subtitle = "The vertical bar at each point shows the standard deviation around the mean. The blue line represents the linear trend.",
    caption = "Source: NSIDC 25km Sea Ice Index dataset"
  ) +
  theme(
    plot.caption = element_text(size = 8, color = "black"),
    plot.margin = unit(c(5.5, 10, 5.5, 5.5), "points"),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black", size = 12, face = "bold"),
    panel.grid = element_blank(),
    axis.line = element_line(color = axis_color),
    legend.position = "none"
  )

print(p)
# 
# # Save the filtered data
# write.csv(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.csv", row.names = FALSE)
# saveRDS(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.rds")
```
Offshore Trends

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(wesanderson)
library(scales)

# Load your dataset
nsidc_metrics <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.rds")

# Add year, month, and day columns
nsidc_metrics <- nsidc_metrics %>%
  mutate(Year = year(Date), Month = month(Date), Day = day(Date))

# Filter for Full_Study_Area and June, July, August, and September
nsidc_metrics_filtered <- nsidc_metrics %>%
  filter(Region == "Offshore_Region" & Month %in% c(6, 7, 8, 9))

# Add a month2 column with month names
nsidc_metrics_filtered <- nsidc_metrics_filtered %>%
  mutate(month2 = month(Date, label = TRUE, abbr = FALSE))

# Choose a light red color for the SD bars
sd_color <- "#7AB8BF"  # Light red color (you can adjust to your preferred shade)
axis_color <- "gray70"  # Light gray color for the axis lines

# Create the plot
p <- nsidc_metrics_filtered %>%
  group_by(month2, Year) %>%
  summarise(mean_extent = mean(IceExtent_km, na.rm = TRUE), 
            sd_extent = sd(IceExtent_km, na.rm = TRUE), .groups = 'drop') %>%
  ggplot(aes(x = Year, y = mean_extent)) +
  geom_errorbar(aes(ymin = mean_extent - sd_extent, ymax = mean_extent + sd_extent), 
                width = 0.2, color = sd_color) +
  geom_point(color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  facet_wrap(~month2, scales = "fixed", ncol = 3) +
  xlab(NULL) +
  ylab(bquote("Ice extent" ~ (km^2 %*% 10^6))) +
  scale_y_continuous(labels = scales::scientific) +
  labs(
    title = sprintf(
      "Offshore Region: Sea ice extent between %d and %d",
      min(nsidc_metrics_filtered$Year),
      max(nsidc_metrics_filtered$Year)
    ),
    subtitle = "The vertical bar at each point shows the standard deviation around the mean. The blue line represents the linear trend.",
    caption = "Source: NSIDC 25km Sea Ice Index dataset"
  ) +
  theme(
    plot.caption = element_text(size = 8, color = "black"),
    plot.margin = unit(c(5.5, 10, 5.5, 5.5), "points"),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black", size = 12, face = "bold"),
    panel.grid = element_blank(),
    axis.line = element_line(color = axis_color),
    legend.position = "none"
  )

print(p)
# 
# # Save the filtered data
# write.csv(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.csv", row.names = FALSE)
# saveRDS(nsidc_metrics_filtered, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics_filtered.rds")

```



## **Deprecated (Non-Parallel) Scripts**


**Calculate Regional Ice Extent Metrics (Not Parallel)**

```{r}

library(terra)
library(lubridate)

# Load data
nsidc_winter <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/NSIDC_25km_Winter_Months.nc")
study_areas <- vect("D:/Manuscripts_localData/FrostBound_AQ/Datasets/gis-layers/study-area/Frostbound_AQ_Subregions_EPSG_3976.shp")

# Define the calculation function
calculate_layer_stats <- function(layer_data, region = NULL) {
  if (month(time(layer_data)) %in% 6:9) {
    if (!is.null(region)) {
      layer_data <- mask(layer_data, region)
    }
    layer_data[values(layer_data) == 0] <- NA
    mean_sic <- as.numeric(global(layer_data, fun = 'mean', na.rm = TRUE))
    cell_area_sq_meters <- prod(res(layer_data))
    valid_ice_cells <- sum(values(layer_data) >= 15, na.rm = TRUE)
    total_ice_area_sq_km <- (valid_ice_cells * cell_area_sq_meters) / 1e6
    return(list(date = time(layer_data), mean_sic = mean_sic, ice_extent_km = total_ice_area_sq_km))
  } else {
    return(NULL)
  }
}

# Process all layers and regions
results <- list()
index <- 1

for (i in 1: nlyr(nsidc_winter)) {
  layer_data <- nsidc_winter[[i]]
  layer_date <- time(layer_data)
  print(paste("Processing layer:", i, "Date:", layer_date))  # Diagnostic print
  
  total_result <- calculate_layer_stats(layer_data)
  if (!is.null(total_result)) {
    total_result$region <- "Total Study Area"
    results[[index]] <- total_result
    index <- index + 1
  }
  
  for (region in unique(study_areas$Region)) {
    region_mask <- study_areas[study_areas$Region == region, ]
    region_result <- calculate_layer_stats(layer_data, region_mask)
    if (!is.null(region_result)) {
      region_result$region <- region
      results[[index]] <- region_result
      index <- index + 1
    }
  }
}

# Combine results and save
result_df <- do.call(rbind, lapply(results, function(x) {
  if (!is.null(x)) {
    data.frame(Date = x$date, MeanSIC = x$mean_sic, IceExtent_km = x$ice_extent_km, Region = x$region)
  } else {
    NULL
  }
}))
write.csv(result_df, output_path, row.names = FALSE)
print("Processing complete. Data saved.")

# Summary of processing
print(paste("Total layers processed:", index - 1))
print(summary(result_df))




```

**Sea Ice Extent Analysis: Sea Ice Extent Generalized Least Squares**

```{r}
# Load necessary libraries
library(dplyr)
library(lubridate)
library(ggplot2)
library(nlme)
library(wesanderson)
library(scales)

# Load your dataset
nsidc_metrics <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.rds")

# Filter for winter months (June - September) and add Year, Month, and Day columns
winter_metrics <- nsidc_metrics %>%
  mutate(Year = year(Date), Month = month(Date), Day = day(Date)) %>%
  filter(Month %in% c(6, 7, 8, 9))

# Calculate monthly averages of ice extent for each region
monthly_averages <- winter_metrics %>%
  group_by(Year, Month, Region) %>%
  summarise(mean_extent = mean(IceExtent_km, na.rm = TRUE), 
            sd_extent = sd(IceExtent_km, na.rm = TRUE), .groups = 'drop') %>%
  ungroup()

# Create a time variable for trend analysis
monthly_averages <- monthly_averages %>%
  mutate(Time = as.numeric(Year) + (Month - 1) / 12,
         month2 = month(ymd(paste(Year, Month, 1)), label = TRUE, abbr = FALSE))

# Define a function to fit a GLS model for a given region and month
fit_gls_model <- function(region_data) {
  gls_model <- gls(mean_extent ~ Time, correlation = corAR1(), data = region_data)
  return(gls_model)
}

# Apply the model for each unique region and month combination and store the results
regions <- unique(monthly_averages$Region)
months <- unique(monthly_averages$Month)
gls_results <- list()

for (region in regions) {
  for (month in months) {
    region_month_data <- monthly_averages %>% filter(Region == region, Month == month)
    if (nrow(region_month_data) > 1) {
      gls_results[[paste(region, month, sep = "_")]] <- fit_gls_model(region_month_data)
    }
  }
}

# Display summary of the GLS model for each region and month
for (key in names(gls_results)) {
  cat("Region and Month:", key, "\n")
  print(summary(gls_results[[key]]))
  cat("\n")
}

# Directory to save the plots
output_directory <- "D:/Manuscripts_localData/FrostBound_AQ/Results/gls_trends"
dir.create(output_directory, recursive = TRUE, showWarnings = FALSE)

# Color settings
sd_color <- "#7AB8BF"  # Light red color (you can adjust to your preferred shade)
axis_color <- "gray70"  # Light gray color for the axis lines

# Visualize the trend for each region and save the plots
for (region in regions) {
  region_data <- monthly_averages %>% filter(Region == region)
  
  p <- region_data %>%
    ggplot(aes(x = Year, y = mean_extent)) +
    geom_errorbar(aes(ymin = mean_extent - sd_extent, ymax = mean_extent + sd_extent), 
                  width = 0.2, color = sd_color) +
    geom_point(color = "black") +
    geom_smooth(method = "lm", color = "blue", se = FALSE) +
    facet_wrap(~month2, scales = "fixed", ncol = 3) +
    xlab(NULL) +
    ylab(bquote("Ice extent" ~ (km^2 %*% 10^6))) +
    scale_y_continuous(labels = scales::scientific) +
    labs(
      title = sprintf(
        "%s: Sea ice extent between %d and %d",
        region,
        min(region_data$Year),
        max(region_data$Year)
      ),
      subtitle = "The vertical bar at each point shows the standard deviation around the mean. The blue line represents the linear trend.",
      caption = "Source: NSIDC 25km Sea Ice Index dataset"
    ) +
    theme(
      plot.caption = element_text(size = 8, color = "black"),
      plot.margin = unit(c(5.5, 10, 5.5, 5.5), "points"),
      panel.background = element_rect(fill = "white"),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(colour = "black", size = 12, face = "bold"),
      panel.grid = element_blank(),
      axis.line = element_line(color = axis_color),
      legend.position = "none"
    )
  
  # Save the plot
  ggsave(filename = paste(output_directory, paste("25km_Extent_trend_plot_", region, ".png", sep = ""), sep = "/"), plot = p)
  

  # Save the plot as EPS
  ggsave(filename = paste(output_directory, paste("25km_Extent_trend_plot_", region, ".eps", sep = ""), sep = "/"), plot = p, device = "eps")
}


# Extract and print coefficients and p-values for each region and month
coefficients_gls <- lapply(gls_results, function(model) summary(model)$tTable)
names(coefficients_gls) <- names(gls_results)

# Print coefficients and p-values
for (key in names(coefficients_gls)) {
  cat("Region and Month:", key, "\n")
  print(coefficients_gls[[key]])
  cat("\n")
}


```



**Sea Ice Concentration Analysis: Generalized Least Squares**

```{r}
# Load necessary libraries
library(dplyr)
library(lubridate)
library(ggplot2)
library(nlme)
library(wesanderson)
library(scales)

# Load your dataset
nsidc_metrics <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/analysis/NSIDC_25km_Daily_Metrics.rds")

# Filter for winter months (June - September) and add Year, Month, and Day columns
winter_metrics <- nsidc_metrics %>%
  mutate(Year = year(Date), Month = month(Date), Day = day(Date)) %>%
  filter(Month %in% c(6, 7, 8, 9))

# Calculate monthly averages of sea ice concentration for each region
monthly_averages <- winter_metrics %>%
  group_by(Year, Month, Region) %>%
  summarise(mean_concentration = mean(MeanSIC, na.rm = TRUE), 
            sd_concentration = sd(MeanSIC, na.rm = TRUE), .groups = 'drop') %>%
  ungroup()

# Create a time variable for trend analysis
monthly_averages <- monthly_averages %>%
  mutate(Time = as.numeric(Year) + (Month - 1) / 12,
         month2 = month(ymd(paste(Year, Month, 1)), label = TRUE, abbr = FALSE))

# Define a function to fit a GLS model for a given region and month
fit_gls_model <- function(region_data) {
  gls_model <- gls(mean_concentration ~ Time, correlation = corAR1(), data = region_data)
  return(gls_model)
}

# Apply the model for each unique region and month combination and store the results
regions <- unique(monthly_averages$Region)
months <- unique(monthly_averages$Month)
gls_results <- list()

for (region in regions) {
  for (month in months) {
    region_month_data <- monthly_averages %>% filter(Region == region, Month == month)
    if (nrow(region_month_data) > 1) {
      gls_results[[paste(region, month, sep = "_")]] <- fit_gls_model(region_month_data)
    }
  }
}

# Display summary of the GLS model for each region and month
for (key in names(gls_results)) {
  cat("Region and Month:", key, "\n")
  print(summary(gls_results[[key]]))
  cat("\n")
}

# Directory to save the plots
output_directory <- "D:/Manuscripts_localData/FrostBound_AQ/Results/gls_trends"
dir.create(output_directory, recursive = TRUE, showWarnings = FALSE)

# Color settings
sd_color <- "#7AB8BF"  # Light red color 
axis_color <- "gray70"  # Light gray color for the axis lines

# Visualize the trend for each region and save the plots
for (region in regions) {
  region_data <- monthly_averages %>% filter(Region == region)
  
  p <- region_data %>%
    ggplot(aes(x = Year, y = mean_concentration)) +
    geom_errorbar(aes(ymin = mean_concentration - sd_concentration, ymax = mean_concentration + sd_concentration), 
                  width = 0.2, color = sd_color) +
    geom_point(color = "black") +
    geom_smooth(method = "lm", color = "blue", se = FALSE) +
    facet_wrap(~month2, scales = "fixed", ncol = 3) +
    xlab(NULL) +
    ylab("Mean Sea Ice Concentration (%)") +
    labs(
      title = sprintf(
        "%s: Sea ice concentration between %d and %d",
        region,
        min(region_data$Year),
        max(region_data$Year)
      ),
      subtitle = "The vertical bar at each point shows the standard deviation around the mean. The blue line represents the linear trend.",
      caption = "Source: NSIDC 25km Sea Ice Index dataset"
    ) +
    theme(
      plot.caption = element_text(size = 8, color = "black"),
      plot.margin = unit(c(5.5, 10, 5.5, 5.5), "points"),
      panel.background = element_rect(fill = "white"),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(colour = "black", size = 12, face = "bold"),
      panel.grid = element_blank(),
      axis.line = element_line(color = axis_color),
      legend.position = "none"
    )
  
  # Save the plot
  ggsave(filename = paste(output_directory, paste("25km_Concentration_trend_plot_", region, ".png", sep = ""), sep = "/"), plot = p)
  
  # Save the plot as EPS
  ggsave(filename = paste(output_directory, paste("25km_Concentration_trend_plot_", region, ".eps", sep = ""), sep = "/"), plot = p, device = "eps")
}

# Extract and print coefficients and p-values for each region and month
coefficients_gls <- lapply(gls_results, function(model) summary(model)$tTable)
names(coefficients_gls) <- names(gls_results)

# Print coefficients and p-values
for (key in names(coefficients_gls)) {
  cat("Region and Month:", key, "\n")
  print(coefficients_gls[[key]])
  cat("\n")
}


```



```{r}
```





```{r}
```


```{r}
```




```{r}
```





```{r}
```





```{r}
```


```{r}
```




```{r}
```





```{r}
```





```{r}
```


```{r}
```




```{r}
```





```{r}
```





```{r}
```


```{r}
```




```{r}
```





```{r}
```





```{r}
```
