---
title: "12.5 km Harmonization"
author: "Michael Wethington"
date: "2024-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**How is this different from the standard pipeline?**

1. All Sea ice concentration values (0-100) area retained. NASA white pages indicate that values less than 15 are unreliable for sea ice concentration. In this case they have been retained in order to have the maximum cell-level daily comparison between AMSR and Sea Ice Index
2. Sea Ice Index Coastline mask is applied to the AMSR dataset in order to ensure proper overlap in cells.  





**Processing Pipeline**
Step 1:
  Create a 12.5 km coastline mask from the NSIDC Sea Ice Index 25 km dataset 

```{r}

# Load the terra library
library(terra)

# Step 1: Load the coastline raster
coastline_raster <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/staged/S_20120704_concentration_v3.0.tif")

# # Step 2: Load the .nc raster
nc_raster <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR-Unified_12km/stack/AMSR-Unified_SIC_Full-Catalog_Time_Series_Subregions.nc")


reclassified_raster <- ifel(coastline_raster == 2530, 2530, NA)

# Step 3: Resample the coastline raster to match the resolution of the .nc raster
resampled_coastline <- resample(reclassified_raster, nc_raster, method="bilinear")

writeRaster(resampled_coastline, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR_Study_Area_Coastline.tif", overwrite=TRUE)
```


#Step 2: 
    Take He5 files
    Reclassify missing data and land sentinel values to NA
    Ensure projection to 3976
    Convert to Geotiff and export to "proccessed



```{r}


# Load required libraries
library(terra)   # For spatial data analysis and raster manipulation
library(rhdf5)   # For reading HDF5 files
library(sf)      # For handling spatial vector data
library(doParallel)
library(foreach)


#Define output directory
output_dir <-  "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/12km_AMSR-Unified/processed-unclipped"

if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}


# Define a function to process a single HDF5 file
process_hdf5_file <- function(h5_file, input_dir, output_dir) {
  # Extract the date from the file path using regex, assuming YYYYMMDD format before ".he5"
  date_pattern <- ".*(\\d{4})(\\d{2})(\\d{2})\\.he5$"
  date <- sub(date_pattern, "\\1-\\2-\\3", basename(h5_file))
  
  # Read Sea Ice Concentration data and apply scale factor
  SeaIce_scaled <- h5read(h5_file, "/HDFEOS/GRIDS/SpPolarGrid12km/Data Fields/SI_12km_SH_ICECON_DAY")
  
  #Reclassify sentinel values
  # Set value ranges based on documentation
  SeaIce_scaled[SeaIce_scaled == 110] <- NA  # Missing data
  SeaIce_scaled[SeaIce_scaled == 120] <- NA  # Land
  
  # Read spatial dimension data for georeferencing
  lat <- h5read(h5_file, "/HDFEOS/GRIDS/SpPolarGrid12km/lat")
  lon <- h5read(h5_file, "/HDFEOS/GRIDS/SpPolarGrid12km/lon")
  
  # Flatten the matrices to vectors
  SeaIce_vec <- as.vector(SeaIce_scaled)
  lat_vec <- as.vector(lat)
  lon_vec <- as.vector(lon)
  
  # Combine vectors into a data frame and convert to an 'sf' object
  sic_points <- data.frame(lon = lon_vec, lat = lat_vec, SIC = SeaIce_vec)
  sic_points_sf <- st_as_sf(sic_points, coords = c("lon", "lat"), crs = 4326, agr = "constant")
  
  # Transform geographic coordinates to the Antarctic Polar Stereographic South projection (EPSG:3412)
  sic_points_sf_transformed <- st_transform(sic_points_sf, crs = "EPSG:3976")
  
  # Create a raster template with dimensions and extent matching the spatial points
  r_corrected <- rast(nrows=664, ncols=632, xmin=-3950000, xmax=3950000, ymin=-3950000, ymax=4350000, crs="EPSG:3412")
  
  # Rasterize the transformed point data onto the raster template
  r_sic <- rasterize(sic_points_sf_transformed, r_corrected, field="SIC", fun=mean)
  
  # Define the reclassification matrix
  # rcl_matrix <- matrix(c(110, NA, 120, NA), ncol = 2, byrow = TRUE)
  # r_sic <- terra::classify(r_sic, rcl_matrix)
  
  # Define the output filename using the output directory and date
  output_filename <- paste0(output_dir, "/AMSR-Unified_SeaIce_12km_", gsub("-", "", date), ".tif")
  
  # Save the GeoTIFF in the output directory
  writeRaster(r_sic, filename=output_filename, overwrite = TRUE)
  
  print(paste0("Raster:", output_filename, " exported"))
  
  # Clear memory of large objects no longer needed
  rm(date, date_pattern, r_sic, sic_points, sic_points_sf, sic_points_sf_transformed, SeaIce_scaled, SeaIce_vec, lat, lat_vec, lon, lon_vec)
  
  # Explicitly call garbage collection to free up memory space
  gc()
}


# Setup parallel environment
numCores <- detectCores()
cl <- makeCluster(numCores - 1)
registerDoParallel(cl)



# Directory containing HDF5 files
input_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR-Unified_12km/staged"



# List all HDF5 files in the directory
hdf5_files <- list.files(input_dir, pattern = "\\.he5$", full.names = TRUE)

# Apply the processing function to each HDF5 file in parallel
results <- foreach(h5_file = hdf5_files, .packages = c("terra", "rhdf5", "sf")) %dopar% {
  process_hdf5_file(h5_file, input_dir, output_dir)
}

# Stop the cluster
stopCluster(cl)
```



Step 3:
  Clip Geotiffs to Study Area using raster mask function


```{r}
# Load required libraries
library(terra)   # For spatial data analysis and raster manipulation

# Define directories
input_raster_dir <-  "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/12km_AMSR-Unified/processed-unclipped"
output_raster_dir <-  "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/12km_AMSR-Unified/processed-clipped"



# Ensure output directory exists
if (!dir.exists(output_raster_dir)) {
  dir.create(output_raster_dir, recursive = TRUE)
}

# Load the study area shapefile
study_area_shp <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/gis-layers/study-area/Frostbound_AQ_Subregions_EPSG_3976.shp"
study_area <- vect(study_area_shp)

# List all raster files in the input directory and sort them by date
raster_files <- list.files(path = input_raster_dir, pattern = "\\.tif$", full.names = TRUE)

# Function to extract date from filename for sorting
extractDate <- function(filename) {
  # Regex to extract date assuming format YYYYMMDD in filename
  dates <- as.Date(sub(".*(\\d{4})(\\d{2})(\\d{2}).*", "\\1-\\2-\\3", basename(filename)), format="%Y-%m-%d")
  return(dates)
}

# Sort files by date
raster_files <- raster_files[order(sapply(raster_files, extractDate))]

# Process each raster file
for (file in raster_files) {
  r <- rast(file)
  # Crop raster to the extent of the study area
  r_cropped <- crop(r, study_area)
  
  # Define the output filename using the original file's name
  output_filename <- file.path(output_raster_dir, basename(file))
  
  # Save the cropped raster to the output directory
  writeRaster(r_cropped, filename=output_filename, overwrite=TRUE)
  
  cat("Processed and saved:", output_filename, "\n")
}

```

Step 4:
  Mask out the NSIDC 25 km Coastline from the AMSR rasters
  
```{r}

# Load the terra library
library(terra)

# Step 1: Load the coastline mask raster
coastline_mask <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR_Study_Area_Coastline.tif")

# Step 2: Load the .nc raster stack
nc_raster_stack <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR-Unified_12km/stack/AMSR-Unified_SIC_Full-Catalog_Time_Series_Subregions.nc")

# Check alignment and resample if necessary
  coastline_mask_resampled <- resample(coastline_mask, nc_raster_stack, method="near")


# Step 3: Create a binary mask (assuming 2530 is the coastline)
binary_mask <- ifel(coastline_mask_resampled == 2530, NA, 1)

# Step 4: Apply the mask to the entire raster stack
masked_nc_raster_stack <- mask(nc_raster_stack, binary_mask, maskvalue=1)

# Step 5: Save the masked raster stack if needed
writeCDF(masked_nc_raster_stack, "D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR-Unified_12km/stack/AMSR-Unified_12km_Harmonization.nc", overwrite=TRUE)

# Print a confirmation
print("The coastline mask has been applied to the entire raster stack and the masked stack has been saved.")


```


Step 4:
#Load Geotiffs into Raster Stack
```{r}
library(terra)
library(lubridate)

# Define the function to extract dates
extractDate <- function(filename) {
  base_name <- basename(filename)
  date_string <- sub(".*_12km_(\\d{8}).*", "\\1", base_name)
  return(as.Date(date_string, format="%Y%m%d"))
}



# Load raster files
raster_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/12km_AMSR-Unified/processed"
raster_files <- list.files(path = raster_dir, pattern = "\\.tif$", full.names = TRUE)
raster_files <- raster_files[order(sapply(raster_files, extractDate))]

# Create a raster stack
raster_stack <- rast(lapply(raster_files, rast))
names(raster_stack) <- sapply(raster_files, function(x) format(extractDate(x), "%Y-%m-%d"))

# Assigning time values
dates <- as.Date(sapply(raster_files, extractDate))
time(raster_stack) <- dates




# Load the NSIDC dataset to use to compare dates
nsidc_stack <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/NSIDC-Sea-Ice-Index/stack/NSIDC-Sea-Ice-Index_25km_SIC_Time_Series_Subregions.nc")

# Extract dates from the time attribute and convert to Date objects
amsr_dates <- as.Date(time(raster_stack))
nsidc_dates <- as.Date(time(nsidc_stack))

# Define winter months: June, July, August, September
winter_months <- c(6, 7, 8, 9)

# Filter dates for winter months
amsr_winter_dates <- amsr_dates[month(amsr_dates) %in% winter_months]
nsidc_winter_dates <- nsidc_dates[month(nsidc_dates) %in% winter_months]

# Subset stacks based on filtered dates
amsr_winter_stack <- raster_stack[[month(amsr_dates) %in% winter_months]]
nsidc_winter_stack <- nsidc_stack[[month(nsidc_dates) %in% winter_months]]

# Check the resulting subsets
print(amsr_winter_stack)
print(nsidc_winter_stack)


# Extract dates from the time attribute and convert to Date objects
amsr_winter_dates <- as.Date(time(amsr_winter_stack))
nsidc_winter_dates <- as.Date(time(nsidc_winter_stack))

# Find dates in NSIDC that are also in AMSR
intersect_dates <- amsr_winter_dates %in% nsidc_winter_dates

# Subset NSIDC stack by intersecting dates
amsr_winter_subset <- amsr_winter_stack[[intersect_dates]]

# Check the resulting subset
print(amsr_winter_subset)



# Save and Winter raster stack
output_path <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/12km_AMSR-Unified/stack/AMSR-Unified_Subregion_Harmonization_Winter.nc"
writeCDF(raster_stack, filename = output_path, overwrite = TRUE)

loaded_stack <- rast(output_path)
print(loaded_stack)
print(time(loaded_stack))


# Assuming 'loaded_stack' is your loaded SpatRaster with supposed time data
time_data <- time(loaded_stack)
if (length(time_data) > 0 && !all(is.na(time_data))) {
  print("Time data are present and appear correctly set.")
  print(head(time_data, 20))  # Print the first 20 time entries to verify
} else {
  print("No valid time data found or all entries are NA.")
}




```









#Generate Raster Products for Annual and Monthly Averages

```{r}
library(terra)
library(sf)

# Load the raster stack
r_fp <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/12km_AMSR-Unified/stack/AMSR-Unified_Continental_Harmonization_Winter.nc"
r_stack <- rast(r_fp)

# # Apply SIC > 15 filter directly on the stack
# r_stack <- ifel(r_stack > 15, r_stack, NA)

# Create a classification matrix to reclassify values
# Values > 15 remain unchanged, <= 15 become NA
# We create a matrix that specifically changes values <= 15 to NA
# We assume SIC values are between 0 and 100, adjust if the scale is different
# rclmat <- matrix(c(0, 15, NA, 16, 100, 16:100), ncol=3, byrow=TRUE)

# Apply classification
# r_classified <- classify(r_stack, rclmat)

# # Ensure the time dimension is available and correctly formatted
# if (is.na(time(r_stack)[1])) {
#   message("Time dimension not properly set. Setting time dimension...")
#   start_date <- as.Date("2012-07-03")  # Adjust according to your actual data
#   time_values <- seq(start_date, length.out = nlyr(r_stack), by = "day")
#   time(r_stack) <- time_values
# }

# Define winter months for the Southern Hemisphere
winter_months <- c("06", "07", "08", "09")

# Prepare containers for results
winter_monthly_means <- list()
annual_winter_means <- list()

# Extract dates, years, and months
dates <- as.Date(time(r_stack))
years <- format(dates, "%Y")
months <- format(dates, "%m")

# Aggregate by winter months for each year
for (year in unique(years)) {
  # Calculate mean for each winter month individually
  for (month in winter_months) {
    idx <- which(years == year & months == month)
    if (length(idx) > 0) {
      winter_monthly_means[[paste(year, month, sep="-")]] <- mean(r_stack[[idx]], na.rm = TRUE)
    }
  }
  
  # Calculate mean for the entire winter season each year
  winter_idx <- which(years == year & months %in% winter_months)
  if (length(winter_idx) > 0) {
    annual_winter_means[[year]] <- mean(r_stack[[winter_idx]], na.rm = TRUE)
  }
}

# Convert lists of rasters to SpatRasters
winter_monthly_mean_raster <- rast(winter_monthly_means)
annual_winter_mean_raster <- rast(annual_winter_means)

# Name layers by year-month for monthly and by year for seasonal
names(winter_monthly_mean_raster) <- names(winter_monthly_means)
names(annual_winter_mean_raster) <- names(annual_winter_means)


# Paths for output NetCDF files
month_raststack_output_path <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/12km_AMSR-Unified/stack/substack/AMSR-Unified_harmonization_winter_monthly_mean_extent.nc"

annual_raststack_output_path <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/12km_AMSR-Unified/stack/substack/AMSR-Unified_harmonization_winter_annual_mean_extent.nc"

# Write results to NetCDF files using writeCDF
writeCDF(winter_monthly_mean_raster, month_raststack_output_path, varname="Winter_Monthly_Mean_SIC", overwrite=TRUE)
writeCDF(annual_winter_mean_raster, annual_raststack_output_path, varname="Winter_Annual_Mean_SIC", overwrite=TRUE)


```


#Plot Annual Raster Trend
```{r}
library(terra)
library(RColorBrewer)

# Set values below 15 to NA
r_stack[r_stack < 15] <- NA

# Create a time vector corresponding to each layer (assuming annual data)
start_year <- 2012  # Adjust according to your data
time_vector <- start_year:(start_year + nlyr(r_stack) - 1)

# Perform regression using only non-NA values
trend_raster <- regress(r_stack, x=time_vector, formula=y~x, na.rm=TRUE)

# Check the results
names(trend_raster)

# Plot the trend slopes (ensure the correct layer is referenced)
diverging_palette <- colorRampPalette(RColorBrewer::brewer.pal(11, "RdBu"))(10)
plot(trend_raster[["x"]], main="Trend of Annual Winter Sea Ice Concentration (Slope)",
     col=rev(diverging_palette),
     breaks=seq(from=-5, to=5, by=0.1), na.color="white")  # Adjust breaks and NA color as needed

# Optionally add a legend
legend("bottomright", legend=c("-5", "0", "5"), fill=c(diverging_palette[1], diverging_palette[50], diverging_palette[100]),
       title="Change per Year", bty="n")

```D:\Manuscripts_localData\FrostBound_AQ\Datasets\dataset-harmonization\12km_AMSR-Unified\stack\AMSR-Unified_Continental_Harmonization_Winter.nc"