---
title: "NSIDC-Sea-Ice-Index_Dataset-Harmonization_Processing-Pipeline"
author: "Michael J. Wethington"
date: "2024-05-10"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Outline: 

Input: NSIDC Sea Ice Index geotiffs

Steps:
    Collect Directory of Geotiffs and collate into a Terra Spatraster Stack
    Assign Names by Layer Date using source geotiff filenames
    Use Dates to Create Time attribute for each layer (for time series analysis)
    Rescale raw data to 1-100 for percentage
    Reclassify sentinel values to NA
    Resample to match AMSR 12.5 km resolution
    Subset raster stack to only winter months
    Note: Input data is EPSG 3976 (same as AMSR-unified 12.5 km)

```{r include=FALSE}
# Load required libraries
library(terra)   # For spatial data analysis and raster manipulation

# Define directories
input_raster_dir <-  "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/25km_Sea-Ice-Index/staged_epsg-3412"
output_raster_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/25km_Sea-Ice-Index/staged_epsg-3976"

# Ensure output directory exists
if (!dir.exists(output_raster_dir)) {
  dir.create(output_raster_dir, recursive = TRUE)
}

# List all raster files in the input directory
raster_files <- list.files(path = input_raster_dir, pattern = "\\.tif$", full.names = TRUE)


# Process each raster file
for (file in raster_files) {
  r <- rast(file)  # Load the raster file
  
  # Reproject the raster to EPSG 3976
  r_reprojected <- project(r, crs("EPSG:3976"))

  # Define the output filename using the original file's name
  output_filename <- file.path(output_raster_dir, basename(file))
  
  # Save the reprojected raster to the output directory
  writeRaster(r_reprojected, filename=output_filename, overwrite=TRUE)
  
  cat("Reprojected to EPSG:3976 and saved:", output_filename, "/n")
}

```



**Step 2** 

    Convert Sentinel values to NA
    Convert files to Geotiff layers

# Set values outside 0-1000 and specific feature values to NA
# Pole hole: 2510
# Coast line: 2530
# Land: 2540
# Missing 2550


```{r} 

library(terra)
library(lubridate)

# Define a function to extract the date from filenames using a regular expression pattern
extractDate <- function(filename) {
  base_name <- basename(filename)
  date_string <- sub("S_(\\d{8})_.*", "\\1", base_name)
  date_converted <- as.Date(date_string, format="%Y%m%d")
  if (is.na(date_converted)) {
    warning("Date conversion failed for: ", base_name)
  }
  return(date_converted)
}



# Set the directory containing the raster files
raster_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/25km_Sea-Ice-Index/staged_epsg-3976"

# List raster files sorted by date extracted from their filenames
raster_files <- list.files(path = raster_dir, pattern = "\\.tif$", full.names = TRUE)
raster_files <- raster_files[order(sapply(raster_files, extractDate))]

# Create a raster stack from the sorted files
raster_stack <- rast(raster_files)

# Assign layer names in the raster stack based on dates extracted from filenames
names(raster_stack) <- sapply(raster_files, function(x) format(extractDate(x), "%Y-%m-%d"))

# Set time attribute for each layer in the stack using extracted dates
dates <- as.Date(sapply(raster_files, extractDate))
time(raster_stack) <- dates


# Set values outside 0-1000 and specific feature values to NA
# Pole hole: 2510
# Coast line: 2530
# Land: 2540
# Missing 2550

# Set specific values representing non-sea ice features to NA in the raster data
values_to_exclude <- c(2510, 2530, 2540, 2550)  # Define non-ice values
raster_stack[raster_stack %in% values_to_exclude] <- NA

# Scale pixel values from a 0-1000 scale to a percentage scale (0-100%)
raster_stack <- raster_stack / 10

# Load the AMSR dataset
amsr_stack <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR-Unified_12km/stack/AMSR-Unified_SIC_Full-Catalog_Time_Series_Subregions.nc")

# Resample the NSIDC dataset to match the resolution of the AMSR dataset using bilinear interpolation
nsidc_resampled <- resample(raster_stack, amsr_stack, method = "bilinear")

# Verify time data in the loaded stack
if (length(time(raster_stack)) > 0 && !all(is.na(time(loaded_stack)))) {
  print("Time data are present and appear correctly set.")
  print(head(time(raster_stack), 20))  # Print the first 20 time entries to verify
} else {
  print("No valid time data found or all entries are NA.")
}

# Subset both datasets for Antarctic winter months (June, July, August, September)
winter_months <- c(6, 7, 8, 9)
amsr_winter_stack <- amsr_stack[[month(time(amsr_stack)) %in% winter_months]]
nsidc_winter_stack <- nsidc_resampled[[month(time(nsidc_resampled)) %in% winter_months]]

# Find and subset by intersecting dates in winter datasets
intersect_dates <- as.Date(time(nsidc_winter_stack)) %in% as.Date(time(amsr_winter_stack))
nsidc_winter_subset <- nsidc_winter_stack[[intersect_dates]]

# Define the output path for the NetCDF file
nc_output_path <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/25km_Sea-Ice-Index/stack/NSIDC-Sea-Ice-Index_25km_Continental_Harmonization.nc"

# Write the subsetted and processed raster stack to a NetCDF file
writeCDF(nsidc_winter_subset, filename = nc_output_path, overwrite = TRUE, varname="sea_ice_concentration", longname="Sea Ice Concentration")


```



**Generatge Monthly/Annual Aggregate Averages**

```{r}

library(terra)
library(sf)

# Load the raster stack
r_fp <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/25km_Sea-Ice-Index/stack/NSIDC-Sea-Ice-Index_25km_Continental_Harmonization_Winter.nc"
r_stack <- rast(r_fp)


# Define winter months for the Southern Hemisphere
winter_months <- c("06", "07", "08", "09")

# Prepare containers for results
winter_monthly_means <- list()
annual_winter_means <- list()

# Extract dates, years, and months
dates <- as.Date(time(r_stack))
years <- format(dates, "%Y")
months <- format(dates, "%m")

# Aggregate by winter months for each year
for (year in unique(years)) {
  # Calculate mean for each winter month individually
  for (month in winter_months) {
    idx <- which(years == year & months == month)
    if (length(idx) > 0) {
      winter_monthly_means[[paste(year, month, sep="-")]] <- mean(r_stack[[idx]], na.rm = TRUE)
    }
  }
  
  # Calculate mean for the entire winter season each year
  winter_idx <- which(years == year & months %in% winter_months)
  if (length(winter_idx) > 0) {
    annual_winter_means[[year]] <- mean(r_stack[[winter_idx]], na.rm = TRUE)
  }
}

# Convert lists of rasters to SpatRasters
winter_monthly_mean_raster <- rast(winter_monthly_means)
annual_winter_mean_raster <- rast(annual_winter_means)

# Name layers by year-month for monthly and by year for seasonal
names(winter_monthly_mean_raster) <- names(winter_monthly_means)
names(annual_winter_mean_raster) <- names(annual_winter_means)


# Paths for output NetCDF files
month_raststack_output_path <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/25km_Sea-Ice-Index/stack/substack/25km_Sea-Ice-Index_harmonization_winter_monthly_mean_extent.nc"

annual_raststack_output_path <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/25km_Sea-Ice-Index/stack/substack/25km_Sea-Ice-Index_harmonization_winter_annual_mean_extent.nc"

# Write results to NetCDF files using writeCDF
writeCDF(winter_monthly_mean_raster, month_raststack_output_path, varname="Winter_Monthly_Mean_SIC", overwrite=TRUE)
writeCDF(annual_winter_mean_raster, annual_raststack_output_path, varname="Winter_Annual_Mean_SIC", overwrite=TRUE)


```




