---
title: "Sea Ice Harmonization Pipeline"
author: "Michael Wethington"
date: "2024-06-28"
output: html_document
---

# Introduction

This tutorial demonstrates how to harmonize two different raster datasets, apply an Artificial Neural Network (ANN) model to predict values, and visualize the results using the `terra`, `nnet`, and `tidymodels` packages in R. The example uses sea ice concentration data from two sources: AMSR-Unified 12.5 km and NSIDC 25 km.

## Overview and Objective

The primary objective of this script is to harmonize and compare sea ice concentration datasets from two different sources, AMSR-Unified 12.5 km and NSIDC 25 km. The script achieves this by performing the following steps:

1. **Load Necessary Libraries**: Import the required R libraries for handling raster data and performing data analysis.
2. **Load Raster Datasets**: Load the sea ice concentration datasets from AMSR-Unified and NSIDC.
3. **Verify and Align Extents and Origins**: Ensure that the spatial extents and origins of the two datasets are aligned to enable direct comparison.
4. **Define and Crop to Intersecting Extent**: Crop both datasets to a common intersecting spatial extent to ensure that they cover the same area.
5. **Resample NSIDC Dataset**: Resample the NSIDC dataset to match the 12.5 km resolution of the AMSR dataset using bilinear interpolation.
6. **Normalize Raster Values**: Normalize the values of both datasets to a 0-1 range for consistent analysis.
7. **Extract Common Dates and Subset Data**: Identify common dates between the two datasets and subset the data to these overlapping periods.
8. **Build and Train ANN Model**: Construct and train an ANN model to analyze the relationship between the AMSR and NSIDC datasets, tuning hyperparameters using grid search.
9. **Apply ANN Model to Raster Stack**: Apply the trained ANN model to predict values in the raster stack, ensuring predictions are clamped between 0 and 1.
10. **Visualize Results**: Plot the original and processed raster layers side by side to visually assess the results.
11. **Model Assessment Using `tidymodels`**: Evaluate the performance of the ANN model using the `tidymodels` package.


```{r}

# Load necessary libraries
if (!require(terra)) install.packages('terra')
if (!require(dplyr)) install.packages('dplyr')
if (!require(reshape2)) install.packages('reshape2')
if (!require(tidyr)) install.packages('tidyr')
if (!require(nnet)) install.packages('nnet')
if (!require(tidymodels)) install.packages('tidymodels')

library(terra)
library(dplyr)
library(reshape2)
library(tidyr)
library(nnet)
library(foreach)
library(doParallel)
library(yardstick)

```


## Train/TestDataset Normalization and Alignment


```{r include=FALSE}

# Function to normalize raster values to a 0-1 range
normalize_raster <- function(r) {
  r_min <- global(r, min, na.rm = TRUE)[[1]]
  r_max <- global(r, max, na.rm = TRUE)[[1]]
  
  if (r_min >= 0 && r_max <= 1) {
    message("Raster values are already within the 0-1 range.")
    return(r)
  } else if (is.finite(r_min) && is.finite(r_max)) {
    normalized_raster <- (r - r_min) / (r_max - r_min)
    return(normalized_raster)
  } else {
    warning("Raster contains only NA values or min/max are not finite.")
    return(r)
  }
}

# Load the AMSR-Unified 12.5 km dataset
print("Loading AMSR-Unified 12.5 km dataset...")
amsr_12km <- tryCatch({
  rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/AMSR-Unified_12km/stack/substack/AMSR_12km_Full_Study_Area.nc")
}, error = function(e) {
  stop("Error loading AMSR-Unified dataset: ", e)
})

# Load the NSIDC 25 km dataset (full timeline from 1979 to 2024)
print("Loading NSIDC 25 km dataset...")
nsidc_25km <- tryCatch({
  rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/25km_Sea-Ice-Index/stack/substack/NSIDC_25km_Full_Study_Area.nc")
}, error = function(e) {
  stop("Error loading NSIDC dataset: ", e)
})

# Verify extents
print("Verifying extents...")
ext_amsr <- ext(amsr_12km)
ext_nsidc <- ext(nsidc_25km)

print(ext_amsr)
print(ext_nsidc)

# Check origins
print("Checking origins...")
origin_amsr <- origin(amsr_12km)
origin_nsidc <- origin(nsidc_25km)

print(origin_amsr)
print(origin_nsidc)

# Align the origin of the NSIDC raster to match the AMSR raster
print("Aligning NSIDC raster to AMSR origin...")
nsidc_aligned <- shift(nsidc_25km, dx = origin_amsr[1] - origin_nsidc[1], dy = origin_amsr[2] - origin_nsidc[2])

# Verify origins again
print("Verifying aligned origins...")
origin_nsidc_aligned <- origin(nsidc_aligned)
print(origin_nsidc_aligned)

# Define the intersecting extent with ymin rounded down to 174999
print("Defining intersecting extent...")
intersection_extent <- ext(
  max(ext_amsr$xmin, ext(nsidc_aligned)$xmin),
  min(ext_amsr$xmax, ext(nsidc_aligned)$xmax),
  174999,
  min(ext_amsr$ymax, ext(nsidc_aligned)$ymax)
)

print(intersection_extent)

# Crop both datasets to the intersecting extent
print("Cropping datasets to intersecting extent...")
amsr_cropped <- crop(amsr_12km, intersection_extent)
nsidc_cropped <- crop(nsidc_aligned, intersection_extent)

# Verify cropped extents
print("Verifying cropped extents...")
ext_amsr_cropped <- ext(amsr_cropped)
ext_nsidc_cropped <- ext(nsidc_cropped)

print(ext_amsr_cropped)
print(ext_nsidc_cropped)

# Resample the cropped NSIDC dataset to 12.5 km resolution using bilinear interpolation
print("Resampling NSIDC dataset to 12.5 km resolution...")
nsidc_resampled_to_12_5km <- resample(nsidc_cropped, amsr_cropped, method = "bilinear")

# Normalize each layer individually and then combine
print("Normalizing AMSR raster values...")
amsr_normalized <- rast(lapply(1:nlyr(amsr_cropped), function(i) {
  print(paste("Normalizing AMSR layer", i))
  normalize_raster(amsr_cropped[[i]])
}))

print("Normalizing NSIDC raster values...")
nsidc_normalized <- rast(lapply(1:nlyr(nsidc_resampled_to_12_5km), function(i) {
  print(paste("Normalizing NSIDC layer", i))
  normalize_raster(nsidc_resampled_to_12_5km[[i]])
}))

# Extract the time information
print("Extracting time information...")
amsr_dates <- time(amsr_normalized)
nsidc_dates <- time(nsidc_normalized)

# Ensure dates are in Date format and find common dates
print("Finding common dates...")
amsr_dates <- as.Date(amsr_dates)
nsidc_dates <- as.Date(nsidc_dates)
common_dates <- as.Date(intersect(amsr_dates, nsidc_dates))

# Find the indices of common dates in the AMSR and NSIDC datasets
print("Finding indices of common dates...")
amsr_indices <- which(amsr_dates %in% common_dates)
nsidc_indices <- which(nsidc_dates %in% common_dates)

# Subset the layers for the overlapping period
print("Subsetting layers for the overlapping period...")
amsr_overlap <- subset(amsr_normalized, amsr_indices)
nsidc_overlap <- subset(nsidc_normalized, nsidc_indices)

# Set the layer names to the corresponding dates
print("Setting layer names for AMSR and NSIDC overlap layers...")
names(amsr_overlap) <- common_dates
names(nsidc_overlap) <- common_dates

# Flatten the rasters to data frames
amsr_df <- as.data.frame(amsr_overlap, xy = TRUE)
nsidc_df <- as.data.frame(nsidc_overlap, xy = TRUE)

# Merge the data frames on the coordinates
merged_df <- merge(amsr_df, nsidc_df, by = c("x", "y"))

# Melt the dataframe to long format
merged_df_long <- melt(merged_df, id.vars = c("x", "y"))

# Separate the date and dataset type from the variable column
merged_df_long <- merged_df_long %>%
  mutate(date = as.Date(gsub("\\..*", "", variable), format = "%Y-%m-%d"),
         dataset = ifelse(grepl("\\.x$", variable), "amsr", "nsidc")) %>%
  select(-variable)

# Spread the dataframe to have separate columns for NSIDC and AMSR
normalized_sea_ice_concentration_df <- merged_df_long %>%
  pivot_wider(names_from = dataset, values_from = value)

# View the structure of the melted and spread dataframe
str(normalized_sea_ice_concentration_df)

# Ensure there are no missing values for the regression
normalized_sea_ice_concentration_df_clean <- normalized_sea_ice_concentration_df %>%
  filter(!is.na(nsidc) & !is.na(amsr))

# Set the output directory
output_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023"
dir.create(output_dir, showWarnings = FALSE)

# Save the dataset
cat("Saving the prepared dataset...\n")
saveRDS(normalized_sea_ice_concentration_df_clean, file = file.path(output_dir, "normalized_sea_ice_concentration_df_clean.rds"))
cat("Prepared dataset saved.\n")

# Save the normalized AMSR and NSIDC raster stacks
cat("Saving the normalized AMSR raster stack...\n")
writeRaster(amsr_normalized, filename = file.path(output_dir, "amsr_normalized_2012-2023.tif"), overwrite = TRUE)
cat("AMSR normalized raster stack saved.\n")

cat("Saving the normalized NSIDC raster stack...\n")
writeRaster(nsidc_normalized, filename = file.path(output_dir, "nsidc_normalized_2012-2023.tif"), overwrite = TRUE)
cat("NSIDC normalized raster stack saved.\n")

# Verify saved datasets
cat("Verifying saved AMSR raster stack...\n")
amsr_normalized_loaded <- rast(file.path(output_dir, "amsr_normalized_2012-2023.tif"))
print(amsr_normalized_loaded)

cat("Verifying saved NSIDC raster stack...\n")
nsidc_normalized_loaded <- rast(file.path(output_dir, "nsidc_normalized_2012-2023.tif"))
print(nsidc_normalized_loaded)

cat("Verifying saved dataframe...\n")
normalized_sea_ice_concentration_df_clean_loaded <- readRDS(file.path(output_dir, "normalized_sea_ice_concentration_df_clean.rds"))
str(normalized_sea_ice_concentration_df_clean_loaded)

range(normalized_sea_ice_concentration_df_clean_loaded$amsr)
range(normalized_sea_ice_concentration_df_clean_loaded$nsidc)

```



**ANN Hyperparameter Tuning (nnet)**

```{r}


# Set the output directory
output_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization"

# Load the saved dataset
cat("Loading the prepared dataset...\n")
merged_df_wide_clean <- readRDS(file = file.path(output_dir, "merged_df_wide_clean.rds"))

# Load the saved raster stack
cat("Loading the NSIDC resampled raster stack...\n")
nsidc_resampled_to_12_5km <- rast(file.path(output_dir, "nsidc_resampled_to_12_5km.tif"))



set.seed(123)  # For reproducibility
train_indices <- sample(1:nrow(merged_df_wide_clean), 0.8 * nrow(merged_df_wide_clean))
train_data <- merged_df_wide_clean[train_indices, ]
test_data <- merged_df_wide_clean[-train_indices, ]


# Set up parallel backend
numCores <- detectCores() - 5
cl <- makeCluster(numCores)
registerDoParallel(cl)

# Define a grid of hyperparameters to search
size_grid <- c(3, 5, 7, 10, 15)
decay_grid <- c(0.01, 0.001, 0.0001)
hyper_grid <- expand.grid(size = size_grid, decay = decay_grid)

# Function to train and evaluate the model
train_and_evaluate <- function(train_data, test_data, size, decay) {
  model <- nnet(amsr ~ nsidc, data = train_data, size = size, linout = TRUE, maxit = 500, decay = decay, trace = FALSE)
  predictions <- predict(model, newdata = test_data)
  mse <- mean((test_data$amsr - predictions)^2)
  return(mse)
}

# Perform grid search in parallel
grid_search_results <- foreach(i = 1:nrow(hyper_grid), .packages = 'nnet', .combine = rbind) %dopar% {
  size <- hyper_grid$size[i]
  decay <- hyper_grid$decay[i]
  mse <- train_and_evaluate(train_data, test_data, size, decay)
  data.frame(size = size, decay = decay, mse = mse)
}

# Stop the cluster
stopCluster(cl)

# Find the best hyperparameters
best_params <- grid_search_results[which.min(grid_search_results$mse), ]
cat("Best hyperparameters: size =", best_params$size, "decay =", best_params$decay, "with MSE =", best_params$mse, "\n")


# Set the output directory
output_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization"

# Save the best hyperparameters and tuning results
cat("Saving the best hyperparameters and tuning results...\n")
saveRDS(best_params, file = file.path(output_dir, "best_hyperparameters.rds"))
saveRDS(grid_search_results, file = file.path(output_dir, "hyperparameter_tuning_results.rds"))
cat("Best hyperparameters and tuning results saved.\n")

```



**Train NNET model**


```{r}


# Load the prepared dataset
cat("Loading the prepared dataset...\n")
normalized_sea_ice_concentration_df_clean <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/normalized_sea_ice_concentration_df_clean.rds")

# Load the saved raster stack
cat("Loading the NSIDC resampled raster stack...\n")
nsidc_resampled_to_12_5km <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/nsidc_normalized_2012-2023.tif")

# Split into training and test sets
set.seed(123)  # For reproducibility
train_indices <- sample(1:nrow(normalized_sea_ice_concentration_df_clean), 0.8 * nrow(normalized_sea_ice_concentration_df_clean))
train_data <- normalized_sea_ice_concentration_df_clean[train_indices, ]
test_data <- normalized_sea_ice_concentration_df_clean[-train_indices, ]

# Train the final ANN model with the best hyperparameters
cat("Training the final ANN model with best hyperparameters...\n")
start_time <- Sys.time()
ann_model <- nnet(amsr ~ nsidc, data = train_data, size = 7, linout = TRUE, decay = 0.001, maxit = 1500, trace = TRUE)
end_time <- Sys.time()
cat("Final ANN training completed in", difftime(end_time, start_time, units = "secs"), "seconds.\n")

# Save the final trained ANN model
cat("Saving the final ANN model...\n")
saveRDS(ann_model, file = "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/final_ann_model_1500-iterations.rds")
cat("Final ANN model saved.\n")

```


##Deploy ANN across the Alligned NSIDC Stack (1979-2023)

**Deploy the ANN**

```{r}

# Load the necessary libraries
library(terra)
library(nnet)
library(dplyr)
library(yardstick)

# Load the trained ANN model
cat("Loading the trained ANN model...\n")
ann_model <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/final_ann_model_1500-iterations.rds")

# Load the NSIDC raster stack
cat("Loading the NSIDC raster stack...\n")
nsidc_stack <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/nsidc_normalized_2012-2023.tif")

# Load the AMSR raster stack
cat("Loading the AMSR raster stack...\n")
amsr_stack <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/amsr_normalized_2012-2023.tif")

# Extract the date information from the raster metadata
nsidc_dates <- as.Date(time(nsidc_stack), origin = "1970-01-01")

# Define the prediction function
predict_harmonized_values <- function(x) {
  df_input <- data.frame(nsidc = x)
  predicted_values <- predict(ann_model, df_input)
  return(predicted_values)
}

# Apply the ANN model to the NSIDC stack
cat("Applying the ANN model to the NSIDC layers...\n")
harmonized_stack <- app(nsidc_stack, predict_harmonized_values)

# Set values above 1 to 1
harmonized_stack <- ifel(harmonized_stack > 1, 1, harmonized_stack)

# Assign the original dates to the harmonized stack
names(harmonized_stack) <- nsidc_dates

# Plot the first layer for verification
par(mfrow = c(1, 2))  # Set up the plotting area to have 1 row and 2 columns
plot(nsidc_stack[[1]], colNA = "red", main = "NSIDC 25km")
plot(harmonized_stack[[1]], colNA = "red", main = "NSIDC 12.5 km Harmonized")

# Set the output directory
# output_dir <- "D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/complete-harmonized-dataset"
# dir.create(output_dir, showWarnings = FALSE)

# # Save the harmonized NSIDC raster stack
# cat("Saving the Harmonized NSIDC raster stack...\n")
# # writeRaster(harmonized_stack, filename = file.path(output_dir, "nsidc_12_5km_normalized_1979-2023.tif"), overwrite = TRUE)
# cat("Harmonized NSIDC raster stack saved.\n")



```

## Assess Model Performance

**ANN Model Assessment**

```{r}

# Load the necessary libraries
library(terra)
library(nnet)
library(dplyr)
library(yardstick)

# Load the trained ANN model
cat("Loading the trained ANN model...\n")
ann_model <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/final_ann_model_1500-iterations.rds")

# Load the train/test dataset
cat("Loading the prepared dataset...\n")
normalized_sea_ice_concentration_df_clean <- readRDS("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/normalized_sea_ice_concentration_df_clean.rds")

# Load the NSIDC raster stack
cat("Loading the NSIDC raster stack...\n")
nsidc_stack <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/nsidc_normalized_2012-2023.tif")

# Load the AMSR raster stack
cat("Loading the AMSR raster stack...\n")
amsr_stack <- rast("D:/Manuscripts_localData/FrostBound_AQ/Datasets/dataset-harmonization/normalized_rasters_2012-2023/amsr_normalized_2012-2023.tif")

# Extract the date information from the raster metadata
nsidc_dates <- as.Date(time(nsidc_stack), origin = "1970-01-01")
amsr_dates <- as.Date(time(amsr_stack), origin = "1970-01-01")

# Find common dates between the two datasets
common_dates <- as.Date(intersect(nsidc_dates, amsr_dates))

# Find indices for common dates in both stacks
nsidc_indices <- which(nsidc_dates %in% common_dates)
amsr_indices <- which(amsr_dates %in% common_dates)

# Subset the layers for the overlapping period
cat("Subsetting layers for the overlapping period...\n")
amsr_overlap <- subset(amsr_stack, amsr_indices)
nsidc_overlap <- subset(nsidc_stack, nsidc_indices)

# Set the layer names to the corresponding dates
cat("Setting layer names for AMSR and NSIDC overlap layers...\n")
names(amsr_overlap) <- common_dates
names(nsidc_overlap) <- common_dates

# Define the prediction function
predict_harmonized_values <- function(x) {
  df_input <- data.frame(nsidc = x)
  predicted_values <- predict(ann_model, df_input)
  return(predicted_values)
}

# Harmonize the NSIDC stack
cat("Applying the ANN model to the NSIDC layers...\n")
harmonized_stack <- app(nsidc_overlap, predict_harmonized_values)

# Compute statistical metrics
compute_metrics <- function(actual, predicted) {
  actual_values <- values(actual)
  predicted_values <- values(predicted)
  # Remove NA values for accurate comparison
  valid_indices <- !is.na(actual_values) & !is.na(predicted_values)
  actual_values <- actual_values[valid_indices]
  predicted_values <- predicted_values[valid_indices]
  
  mae <- yardstick::mae_vec(actual_values, predicted_values)
  rmse <- yardstick::rmse_vec(actual_values, predicted_values)
  r2 <- cor(actual_values, predicted_values)^2
  
  return(list(MAE = mae, RMSE = rmse, R2 = r2))
}

# Split into training and test sets
set.seed(123)  # For reproducibility
train_indices <- sample(1:nrow(normalized_sea_ice_concentration_df_clean), 0.8 * nrow(normalized_sea_ice_concentration_df_clean))
train_data <- normalized_sea_ice_concentration_df_clean[train_indices, ]
test_data <- normalized_sea_ice_concentration_df_clean[-train_indices, ]

# Apply metrics to each layer
metrics_list <- lapply(1:nlyr(amsr_overlap), function(i) {
  actual <- amsr_overlap[[i]]
  predicted <- harmonized_stack[[i]]
  compute_metrics(actual, predicted)
})

# Print out metrics for each layer
for (i in 1:length(metrics_list)) {
  cat("Layer", i, ":\n")
  print(metrics_list[[i]])
}

# Extract individual metric values and calculate the averages
mae_values <- sapply(metrics_list, function(x) x$MAE)
rmse_values <- sapply(metrics_list, function(x) x$RMSE)
r2_values <- sapply(metrics_list, function(x) x$R2)

average_mae <- mean(mae_values, na.rm = TRUE)
average_rmse <- mean(rmse_values, na.rm = TRUE)
average_r2 <- mean(r2_values, na.rm = TRUE)

# Print out average metrics
cat("Average MAE:", average_mae, "\n")
cat("Average RMSE:", average_rmse, "\n")
cat("Average R2:", average_r2, "\n")

# Evaluate the model using yardstick on the test dataset
cat("Evaluating the model on the test dataset...\n")
test_data <- test_data %>%
  mutate(predictions = as.numeric(predict(ann_model, newdata = test_data)))

mae_result <- yardstick::mae(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)
rmse_result <- yardstick::rmse(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)
rsq_result <- yardstick::rsq(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)
smape_result <- yardstick::smape(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)
mase_result <- yardstick::mase(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)
ccc_result <- yardstick::ccc(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)
huber_loss_result <- yardstick::huber_loss(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)
rpiq_result <- yardstick::rpiq(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)
rpd_result <- yardstick::rpd(test_data, truth = amsr, estimate = predictions, na_rm = TRUE)

# Print yardstick metrics
cat("Test Data Metrics:\n")
cat("MAE:", mae_result$.estimate, "\n")
cat("RMSE:", rmse_result$.estimate, "\n")
cat("R2:", rsq_result$.estimate, "\n")
cat("SMAPE:", smape_result$.estimate, "\n")
cat("MASE:", mase_result$.estimate, "\n")
cat("CCC:", ccc_result$.estimate, "\n")
cat("Huber Loss:", huber_loss_result$.estimate, "\n")
cat("RPIQ:", rpiq_result$.estimate, "\n")
cat("RPD:", rpd_result$.estimate, "\n")

# Optionally, plot the first layer of each stack for verification
par(mfrow = c(1, 3))  # Set up the plotting area to have 1 row and 3 columns
plot(nsidc_overlap[[1]], colNA = "red", main = "NSIDC 25km")
plot(harmonized_stack[[1]], colNA = "red", main = "NSIDC 12.5 km Harmonized")
plot(amsr_overlap[[1]], colNA = "red", main = "AMSR Unified 12.5 km")


```












