---
title: "Gentoo Model Parameter Abundance Model"
author: "Michael J. Wethington"
date: "2024-06-04"
output: html_document
---


zi = latent nest abundance (mean-adjusted)

lz = logged abundance (re-expression of above):  lzi,t=log(zi,t). for the ith site in the tth year,

ri = intrinsic growth rate

lp = predicted population growth rate multiplier
la = actual population growth rate multiplier 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(coda)
library(mapppdr)
library(tidyverse)
library(patchwork)
library(leaflet)
library(CCAMLRGIS)
library(rjags)
library(MCMCvis)
library(parallel)
library(stringi)
library(pander)
library(testthat)
library(stringi)

# required_packages <- c("rlang", "fastmap", "digest", "fs", "cachem", "vctrs", "stringi", "glue", "cli", "utf8", "fansi", "dplyr")
# 
# for(pkg in required_packages){
#   if(!require(pkg, character.only = TRUE)){
#     install.packages(pkg)
#   }
# }

```




**Construct Presence-Absence Assumptions csv for the JAGS model**

```{r}

# Parameters
min_season <- 1970
max_season <- 2023
species <- "GEPE"

# Load penguin observations data
penguin_obs <- mapppdr::penguin_obs

# Process the penguin observations dataset to match the required format
penguin_obs_processed <- penguin_obs %>%
  filter(species_id == species) %>%
  mutate(
    presence = ifelse(!is.na(count), 1, 0),
    known_w = 1) %>%
  select(site_id, season, presence, known_w, count, accuracy, type)

# Generate the presence/absence assumptions dataset
presence_absence_assumptions <- expand.grid(
  site_id = unique(penguin_obs$site_id),
  season = min_season:max_season
) %>%
  left_join(penguin_obs_processed, by = c("site_id", "season")) %>%
  mutate(
    presence = ifelse(is.na(presence), 0, presence),
    known_w = ifelse(is.na(known_w), 0, known_w),
    count = ifelse(is.na(count), 0, count)) %>%
  group_by(site_id, season) %>%
  arrange(desc(type), desc(accuracy), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  select(site_id, season, presence, known_w, count, accuracy)

# Visualize presence/absence data
plot_data <- presence_absence_assumptions %>%
  mutate(
    presence = as.factor(presence),
    known_w = as.factor(known_w)
  )

# # Plot with facetting
# ggplot(plot_data, aes(x = factor(season), y = factor(site_id))) +
#   geom_tile(aes(fill = presence, color = known_w), size = 1.5) +
#   scale_fill_manual(values = c("0" = "gray90", "1" = "blue")) +
#   scale_color_manual(values = c("0" = NA, "1" = "black")) +
#   labs(
#     x = "Season",
#     y = "Site ID",
#     fill = "Presence",
#     color = "Known"
#   ) +
#   theme_minimal() +
#   theme(
#     axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
#     axis.text.y = element_text(size = 5),
#     panel.grid = element_blank()
#   ) +
#   facet_wrap(~ site_id, scales = "free_y", ncol = 5) +
#   coord_fixed(ratio = 1)  # Adjust aspect ratio to make tiles more square

# ggsave("gentoo_presence_absence_plot_facetted.png", width = 20, height = 15)



```



**Load the Jags MCMC Output File**

```{r}
# Load the RDA file
load("D:/Manuscripts_localData/FrostBound_AQ/Results/gentoo-abundance-model/model_data_rinits_output.rda")

# Ensure the data is loaded
print(ls())

# Check the structure of the dataset
str(model_data_rinits_output)

```


**Extract the Gentoo Abundance Estimates (lz)**

lza = logged latent seasonal abudance

The logged latent seasonal nest abundance for the ith site in the tth year, while allowing for site extirpation and colonization events, as lzai,t=lzi,t⋅wi,t. Please see #3 in Section 2.5 for the details regarding wi,t.


population growth rate multiplier

```{r}
#cast the mcmc model into a matrix
model_samples <- as.matrix(model_data_rinits_output)


# Filter columns that are logged latent abundance (lz) parameters
lz_columns <- grep("^lz\\[", colnames(model_samples)) #logged abundance
lz_samples <- model_samples[, lz_columns]

#Convert log-abundances to actual abundances
abundance_samples <- exp(lz_samples)

# Summarize the actual abundances
abundance_summary <- apply(abundance_samples, 2, function(x) {
  c(mean = mean(x), median = median(x), 
    lower_95 = quantile(x, 0.025), upper_95 = quantile(x, 0.975))
})

# Convert to a readable data frame (use t to transpose)
abundance_summary_df <- as.data.frame(t(abundance_summary))


# Updated extract_indices function
extract_indices <- function(colname) {
  indices <- gsub("[^0-9,]", "", colname)
  as.integer(unlist(strsplit(indices, ",")))
}

indices <- lapply(colnames(abundance_samples), extract_indices)
sites <- sapply(indices, `[`, 1)
seasons <- sapply(indices, `[`, 2)

abundance_summary_df$site <- sites
abundance_summary_df$season <- seasons

abundance_summary_df <- abundance_summary_df %>% 
  rename(mean_abundance = mean,
         median_mean_abundance = median,
         lower_95_mean_abundance = "lower_95.2.5%",
         upper_95_mean_abundance = "upper_95.97.5%")

head(abundance_summary_df)


```


```{r}
SiteList <- mapppdr::penguin_obs %>%
  # keep all sites that have at least 1 count between min and max season
  dplyr::filter(count > 0 & species_id == species & season >= min_season & season <= max_season) %>%
  # create relative season index 
  mutate(season_relative = season - min_season + 1) %>%
  # determine first season a count is observed for each site
  group_by(site_id) %>%
  summarise(initial_season = min(season_relative)) %>%
  ungroup() %>%
  # join to get other site specific covariates for visualization purposes
  left_join(mapppdr::sites, by = "site_id") %>%
  # create site index for model and visualization
  mutate(site = as.numeric(as.factor(site_id))) %>%
  dplyr::select(site_id, site_name, ccamlr_id, site, initial_season, latitude, longitude)

(n_sites <- nrow(SiteList))

SiteList %>% 
  distinct(site, site_id, latitude, longitude )
  

# Join Site_id Info and Rename
final_data <- left_join(SiteList, abundance_summary_df, by = "site" )

# Ensure final_data is sorted by site and season
final_data <- final_data %>%
  arrange(site, season)

```

**Calculate Growth Rate**

```{r}

# Calculate the growth rate and append it to the data
final_data <- final_data %>%
  group_by(site) %>%
  mutate(growth_rate = mean_abundance / lag(mean_abundance)) %>%
  ungroup()

# Print the head of the final data with growth rates
print(head(final_data))
```


```{r}


```


```{r}

library(stringi)
library(MCMCvis)
library(dplyr)
library(ggplot2)
library(tidyr)

# Load the RDA file
load("D:/Manuscripts_localData/FrostBound_AQ/Results/gentoo-abundance-model/model_data_rinits_output.rda")

# Summarize the latent abundance (lz) parameters
lz_summary <- MCMCsummary(model_data_rinits_output, params = "lz", HPD = TRUE, hpd_prob = .95, round = 3)
print(head(lz_summary))

# Extract samples for lz
lz_samples <- as.matrix(model_data_rinits_output)

# Filter columns that are latent abundance (lz) parameters
lz_columns <- grep("^lz\\[", colnames(lz_samples))
lz_samples <- lz_samples[, lz_columns]

# Convert log-abundances to actual abundances
abundance_samples <- exp(lz_samples)

# Summarize the actual abundances
abundance_summary <- apply(abundance_samples, 2, function(x) {
  c(mean = mean(x), median = median(x), 
    lower_95 = quantile(x, 0.025), upper_95 = quantile(x, 0.975))
})

# Convert to a readable data frame
abundance_summary_df <- as.data.frame(t(abundance_summary))
print(head(abundance_summary_df))

# Extract site and season info from the indices
extract_indices <- function(colname) {
  indices <- gsub("[^0-9,]", "", colname)
  as.integer(unlist(strsplit(indices, ",")))
}

indices <- lapply(colnames(abundance_samples), extract_indices)
sites <- sapply(indices, `[`, 1)
seasons <- sapply(indices, `[`, 2)

abundance_summary_df$site <- sites
abundance_summary_df$season <- seasons

abundance_summary_df <- abundance_summary_df %>% 
  rename(mean_abundance = mean,
         median_abundance = median,
         lower_95_abundance = "lower_95.2.5%",
         upper_95_abundance = "upper_95.97.5%")

head(abundance_summary_df)

# Load and prepare SiteList
SiteList <- mapppdr::penguin_obs %>%
  filter(count > 0 & species_id == species & season >= min_season & season <= max_season) %>%
  mutate(season_relative = season - min_season + 1) %>%
  group_by(site_id) %>%
  summarise(initial_season = min(season_relative)) %>%
  ungroup() %>%
  left_join(mapppdr::sites, by = "site_id") %>%
  mutate(site = as.numeric(as.factor(site_id))) %>%
  select(site_id, site_name, ccamlr_id, site, initial_season, latitude, longitude)

(n_sites <- nrow(SiteList))

# Ensure distinct site and site_id combinations
SiteList %>% 
  distinct(site, site_id)

# Join SiteList with abundance_summary_df
final_data <- left_join(SiteList, abundance_summary_df, by = "site")

# Ensure final_data is sorted by site and season
final_data <- final_data %>%
  arrange(site, season)

# Calculate the growth rate and append it to the data
final_data <- final_data %>%
  group_by(site) %>%
  mutate(growth_rate = mean_abundance / lag(mean_abundance)) %>%
  ungroup()

# Print the head of the final data with growth rates
print(head(final_data))

```


```{r}

# Load required libraries
library(tidyverse)
library(coda)
library(mapppdr)
library(patchwork)
library(leaflet)
library(CCAMLRGIS)
library(rjags)
library(MCMCvis)
library(parallel)
library(stringi)
library(pander)
library(testthat)

# Define parameters
min_season <- 1970
max_season <- 2023
species <- "GEPE"

# Construct Presence-Absence Assumptions CSV for the JAGS model
penguin_obs <- mapppdr::penguin_obs

penguin_obs_processed <- penguin_obs %>%
  filter(species_id == species) %>%
  mutate(
    presence = ifelse(!is.na(count), 1, 0),
    known_w = 1) %>%
  select(site_id, season, presence, known_w, count, accuracy, type)

presence_absence_assumptions <- expand.grid(
  site_id = unique(penguin_obs$site_id),
  season = min_season:max_season
) %>%
  left_join(penguin_obs_processed, by = c("site_id", "season")) %>%
  mutate(
    presence = ifelse(is.na(presence), 0, presence),
    known_w = ifelse(is.na(known_w), 0, known_w),
    count = ifelse(is.na(count), 0, count)) %>%
  group_by(site_id, season) %>%
  arrange(desc(type), desc(accuracy), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  select(site_id, season, presence, known_w, count, accuracy)

# Load the JAGS MCMC Output File
load("D:/Manuscripts_localData/FrostBound_AQ/Results/gentoo-abundance-model/model_data_rinits_output.rda")

# Extract the Gentoo Abundance Estimates (lz)
model_samples <- as.matrix(model_data_rinits_output)

# Filter columns that are logged latent abundance (lz) parameters
lz_columns <- grep("^lz\\[", colnames(model_samples))
lz_samples <- model_samples[, lz_columns]

# Convert log-abundances to actual abundances
abundance_samples <- exp(lz_samples)

# Summarize the actual abundances
abundance_summary <- apply(abundance_samples, 2, function(x) {
  c(mean = mean(x), median = median(x), 
    lower_95 = quantile(x, 0.025), upper_95 = quantile(x, 0.975))
})

# Convert to a readable data frame (use t to transpose)
abundance_summary_df <- as.data.frame(t(abundance_summary))

# Extract site and season info from the indices
extract_indices <- function(colname) {
  indices <- gsub("[^0-9,]", "", colname)
  as.integer(unlist(strsplit(indices, ",")))
}

indices <- lapply(colnames(abundance_samples), extract_indices)
sites <- sapply(indices, `[`, 1)
seasons <- sapply(indices, `[`, 2)

abundance_summary_df$site <- sites
abundance_summary_df$season <- seasons

abundance_summary_df <- abundance_summary_df %>% 
  rename(mean_abundance = mean,
         median_abundance = median,
         lower_95_abundance = "lower_95.2.5%",
         upper_95_abundance = "upper_95.97.5%")

head(abundance_summary_df)

# Load and prepare SiteList
SiteList <- mapppdr::penguin_obs %>%
  filter(count > 0 & species_id == species & season >= min_season & season <= max_season) %>%
  mutate(season_relative = season - min_season + 1) %>%
  group_by(site_id) %>%
  summarise(initial_season = min(season_relative)) %>%
  ungroup() %>%
  left_join(mapppdr::sites, by = "site_id") %>%
  mutate(site = as.numeric(as.factor(site_id))) %>%
  select(site_id, site_name, ccamlr_id, site, initial_season, latitude, longitude)

(n_sites <- nrow(SiteList))

SiteList %>% 
  distinct(site, site_id, latitude, longitude)

# Join SiteList with abundance_summary_df
final_data <- left_join(SiteList, abundance_summary_df, by = "site")

# Ensure final_data is sorted by site and season
final_data <- final_data %>%
  arrange(site, season)

# Calculate the growth rate and append it to the data
final_data <- final_data %>%
  group_by(site) %>%
  mutate(growth_rate = mean_abundance / lag(mean_abundance)) %>%
  ungroup()

# Print the head of the final data with growth rates
print(head(final_data))

```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


