---
title: "Gentoo Model Parameter Abundance Model"
author: "Michael J. Wethington"
date: "2024-06-04"
output: html_document
---


zi = latent nest abundance (mean-adjusted)

lz = logged abundance (re-expression of above):  lzi,t=log(zi,t). for the ith site in the tth year,

ri = intrinsic growth rate

lp = predicted population growth rate multiplier
la = actual population growth rate multiplier 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(coda)
library(mapppdr)
library(tidyverse)
library(patchwork)
library(leaflet)
library(CCAMLRGIS)
library(rjags)
library(MCMCvis)
library(parallel)
library(stringi)
library(pander)
library(testthat)
library(stringi)

# required_packages <- c("rlang", "fastmap", "digest", "fs", "cachem", "vctrs", "stringi", "glue", "cli", "utf8", "fansi", "dplyr")
# 
# for(pkg in required_packages){
#   if(!require(pkg, character.only = TRUE)){
#     install.packages(pkg)
#   }
# }

```




**Construct Presence-Absence Assumptions csv for the JAGS model**

```{r}

# Parameters
min_season <- 1970
max_season <- 2023
species <- "GEPE"

# Load penguin observations data
penguin_obs <- mapppdr::penguin_obs

# Process the penguin observations dataset to match the required format
penguin_obs_processed <- penguin_obs %>%
  filter(species_id == species) %>%
  mutate(
    presence = ifelse(!is.na(count), 1, 0),
    known_w = 1) %>%
  select(site_id, season, presence, known_w, count, accuracy, type)

# Generate the presence/absence assumptions dataset
presence_absence_assumptions <- expand.grid(
  site_id = unique(penguin_obs$site_id),
  season = min_season:max_season
) %>%
  left_join(penguin_obs_processed, by = c("site_id", "season")) %>%
  mutate(
    presence = ifelse(is.na(presence), 0, presence),
    known_w = ifelse(is.na(known_w), 0, known_w),
    count = ifelse(is.na(count), 0, count)) %>%
  group_by(site_id, season) %>%
  arrange(desc(type), desc(accuracy), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  select(site_id, season, presence, known_w, count, accuracy)

# Visualize presence/absence data
plot_data <- presence_absence_assumptions %>%
  mutate(
    presence = as.factor(presence),
    known_w = as.factor(known_w)
  )

# Plot with facetting
ggplot(plot_data, aes(x = factor(season), y = factor(site_id))) +
  geom_tile(aes(fill = presence, color = known_w), size = 1.5) +
  scale_fill_manual(values = c("0" = "gray90", "1" = "blue")) +
  scale_color_manual(values = c("0" = NA, "1" = "black")) +
  labs(
    x = "Season",
    y = "Site ID",
    fill = "Presence",
    color = "Known"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
    axis.text.y = element_text(size = 5),
    panel.grid = element_blank()
  ) +
  facet_wrap(~ site_id, scales = "free_y", ncol = 5) +
  coord_fixed(ratio = 1)  # Adjust aspect ratio to make tiles more square

# ggsave("gentoo_presence_absence_plot_facetted.png", width = 20, height = 15)



```






**Load the Jags MCMC Output File**

```{r}
# Load the RDA file
load("D:/Manuscripts_localData/FrostBound_AQ/Results/gentoo-abundance-model/model_data_rinits_output.rda")

# Ensure the data is loaded
print(ls())

# Check the structure of the dataset
str(model_data_rinits_output)

```


**Extract the Gentoo Abundance Estimates (lz)**

lza = logged latent seasonal abudance

The logged latent seasonal nest abundance for the ith site in the tth year, while allowing for site extirpation and colonization events, as lzai,t=lzi,t⋅wi,t. Please see #3 in Section 2.5 for the details regarding wi,t.


population growth rate multiplier

```{r}
#cast the mcmc model into a matrix
model_samples <- as.matrix(model_data_rinits_output)


# Filter columns that are logged latent abundance (lz) parameters
lz_columns <- grep("^lz\\[", colnames(model_samples)) #logged abundance
lz_samples <- lz_samples[, lz_columns]

#Convert log-abundances to actual abundances
abundance_samples <- exp(lz_samples)

# Summarize the actual abundances
abundance_summary <- apply(abundance_samples, 2, function(x) {
  c(mean = mean(x), median = median(x), 
    lower_95 = quantile(x, 0.025), upper_95 = quantile(x, 0.975))
})

# Convert to a readable data frame (use t to transpose)
abundance_summary_df <- as.data.frame(t(abundance_summary))


```

**Extract site and season info from the indices**
```{r}
# Updated extract_indices function
extract_indices <- function(colname) {
  indices <- gsub("[^0-9,]", "", colname)
  as.integer(unlist(strsplit(indices, ",")))
}

indices <- lapply(colnames(abundance_samples), extract_indices)
sites <- sapply(indices, `[`, 1)
seasons <- sapply(indices, `[`, 2)

abundance_summary_df$site <- sites
abundance_summary_df$season <- seasons
head(abundance_summary_df)


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


